---
title: "Ethnic Population Projection for Birmingham"
author: "Chung Au-Yeung"
csl   : "csl/apa.csl"
bibliography: "references.bib"
date: today
format:
  html:
    include-in-header:
      text: |
        <style>
        .csl-entry a {
          word-break: break-all;
          overflow-wrap: anywhere;
        }
        </style>
    toc: true
    toc-location: left
    number-sections: true
    code-fold: true
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
crossref:
  chapters: true
  sec-prefix: "Section"
  tbl-title: "Table"
  fig-title: "Figure"
editor: 
  markdown: 
    wrap: 72
---

```{r setup}
#| context: setup
#| warning: false
#| message: false

library(tidyverse)
library(readr)
library(readxl)
library(bcctheme)
library(scales)
library(patchwork)
library(here)
library(gt)
```

\newpage

# Introduction

Sub-national population projections disaggregated by ethnic group are
valuable insights to inform local strategic planning, health service
commissioning for local authorities. However, this type of projection is
rare due to its complexity and data availability. A substantial body of
work on ethnic population projection has been developed by researchers
at the University of Leeds, starting with the bi-regional
cohort-component framework set out by @wohland2010ethnic. It was then
revised and extended through the NEWETHPOP project [@rees2015newethpop;
@rees2017newethpop]. Birmingham has one of the youngest and most
ethnically diverse populations in the UK and it is crucial to understand
the compositional change of different ethnicity to anticipate the demand
on public health services in the coming

The purpose of this work is to employ the framework of
@wohland2010ethnic as the primary methodological anchor to produce an
ethnic population projections for Birmingham. The projections are
intended to support short and long-term analyses and planning where
changes in the size and age structure of ethnic populations are
relevant.

@sec-framework describes the overall model specification and
@sec-projectionmodel presents the equations of the projection model.
@sec-data displays all data sources used. @sec-fertility through
@sec-internal-migration provide detailed procedure of processing and
calculating each probability being used in the model including
mortality, fertility, international migration and internal migration.
@sec-assumptions lists out a summary table of all assumptions being
imposed for the projection.

\newpage

# Model framework {#sec-framework}

## Bi-regional structure

This analysis employs the bi-regional cohort component model, treating
Birmingham as the zone of interest and all other local authorities as a
single compainion zone namely the Rest of England and Wales (REW). This
follows the bi-regional design closely, in which each zone of interest
is projected against the REW rather than projecting every other
individual local authority simultaneously [@wohland2010ethnic].
Birmingham is projected fully through the cohort-component accounting
steps described in @sec-projectionmodel . However, REW is *not* obtained
within the model through projecting every pair of zone of interest and
the REW. Instead, REW population by age, sex and ethnic group is served
as an exogenous source obtained and derived from the Office for National
Statistics (ONS) Sub-national Population Projections (SNPP) variant data
. This design choice is justified because the authors does not have data
for all other local authorities and it is beyond the scope of this
project to also produce projections for other local authorities as well.
Hence, the model is a simplified version of the bi-regional model set
out by @wohland2010ethnic.

## Ethnic group classifications

Ethnicity classifications has not been recorded consistently in UK
official statistics over time. Within England and Wales, the census
ethnic categories have changed across the 1991,2001,2011 and 2021
rounds, moving from 9 categories to 16 and 20 in 2021. This implies that
ethnicity recording changes to reflect the shift and evolution of ethnic
groups and population structure. Broad ethnic grouping such as White,
Mixed, Black, Asian and Others are useful for general reporting but they
can hide important demographic differences between groups especially
fertility, mortality and migration patterns. Therefore, this project
uses the 12 harmonised ethnic groups adopted in the later NEWETHPOP
project [@rees2017newethpop]: White British/Irish/Gypsy/Traveller (WBI),
White Other (WHO), Mixed/Multiple Ethnic Groups (MIX), Indian (IND),
Pakistani (PAK), Bangladeshi (BAN), Chinese (CHI), Other Asian (OAS),
Black African (BLA), Black Caribbean (BLC), Other Black (OBL), and Other
Ethnic Group (OTH). This classification provide sufficient demographic
details while allowing ethnic group to be treated consistently in the
projection model. The following @tbl-ethnic-groups shows the ethnicity
mapping to the 12 harmonised groups from different data.

| NewEthPop Code | NewEthPop Label       | Death Register Categories                                                                       | Census 2021 20 Categories                                                                                                                   | MSDS ethnicity coding |
|---------------|---------------|---------------|---------------|---------------|
| WBI            | White British & Irish | British; Irish                                                                                  | White: English, Welsh, Scottish, Northern Irish or British; White: Irish; White: Gypsy or Irish Traveller; White: Roma                      | A, B                  |
| WHO            | White Other           | Any other white background                                                                      | White: Other White                                                                                                                          | C                     |
| MIX            | Mixed                 | White and Black Caribbean; White and Black African; White and Asian; Any other mixed background | Mixed or Multiple ethnic groups: White and Black Caribbean; White and Black African; White and Asian; Other Mixed or Multiple ethnic groups | D, E, F, G            |
| IND            | Indian                | Indian                                                                                          | Asian: Indian                                                                                                                               | H                     |
| PAK            | Pakistani             | Pakistani                                                                                       | Asian: Pakistani                                                                                                                            | J                     |
| BAN            | Bangladeshi           | Bangladeshi                                                                                     | Asian: Bangladeshi                                                                                                                          | K                     |
| CHI            | Chinese               | Chinese                                                                                         | Asian: Chinese                                                                                                                              | R                     |
| OAS            | Other Asian           | Any other Asian background                                                                      | Asian: Other Asian                                                                                                                          | L                     |
| BLA            | Black African         | African                                                                                         | Black: African                                                                                                                              | N                     |
| BLC            | Black Caribbean       | Caribbean                                                                                       | Black: Caribbean                                                                                                                            | M                     |
| OBL            | Other Black           | Any other Black background                                                                      | Black: Other Black                                                                                                                          | P                     |
| OTH            | Other Ethnic Group    | Any other ethnic group                                                                          | Other ethnic group: Arab; Other ethnic group: Any other ethnic group                                                                        | S                     |

: NEWETHPOP ethnic group classification {#tbl-ethnic-groups}

## Ages

This project uses single year of age as the main age classification.
This provides flexibility for reporting and also for further grouping
into wider bands if needed. Given that the estimations of fertility,
mortality and migration rates for every single year of age by ethnicity
and sex can be unstable due to small numbers. The procedure of handling
this issue is explained in later section. The projected age range is
between 0 and 100+ where ages above 100 years are grouped to 100+.

## Projection horizon

The projection is anchored to the 2021 Census population, which provides
the base population by single year of age, sex and ethnic group. The
reference date of this data is not the same as the 2021 mid-year
estimate reference point, the resulting total resident population shows
only small differences. Therefore, it is directly used as the 2021
mid-year estimate. Annual projections are then generated from 2022 to
2047, with the first projected population representing the 2022 mid-year
estimated resident population after accounting the components between
2021 and 2022.

## Scenarios for projections

Three projection scenarios were produced to compare and examine the
sensitivity mainly on the international emigration. The benchmark
emigration flow projection (B-EF) directly uses emigration flow, meaning
that a fixed number for each projection year is being taken away from
the model. The benchmark emigration rates projection (B-ER2022) uses the
emigration flow in the first projection year to clacualte the ethnic,
sex, age emigration rates and are held constant throughout the
projection year. This allow the number of emigrants to vary to changes
in size of the population at risk. The Long-term emigration rates
projection (LT-ER2030) retains the annual emigration-flow assumptions up
to 2029. The ethnic, sex, age emigration rates are then calculated using
the 2030 emigration flows and population at risk and held constant from
2030 to 2047. This allow rates to be based on longer-term emigration
assumptions rather than the initial projection year.

# The projection model {#sec-projectionmodel}

The Birmingham ethnic population projection uses the cohort-component
framework based on the approach employed by @wohland2010ethnic.
Population change is modelled annually through six demographic
components: births, deaths, internal in-migration, internal
out-migration, international immigration and emigration. Since the full
methodology is documented in that working paper, only the key equations
are displayed here. Intermediate derivations and the full set of
underlying equations can be found in the original paper.

## Population accounting

The population at the end of the interval for each ethnic group $e$, age
$x$ and gender $s$ in zone $i$ is given by:

$$
P^{end,i}_{e,s,x,t}
=
P^{start,i}_{e,s,x,t}
-
D^{i}_{e,s,x,t}
-
E^{i}_{e,s,x,t}
-
M^{ir}_{e,s,x,t}
+
M^{ri}_{e,s,x,t}
+
I^{i}_{e,s,x,t}
$$ {#eq-projection-accounting}

the $P^{start,i}_{e,s,x,t}$ is the start population of zone $i$. From
this population, deaths during the interval $D^{i}_{e,s,x,t}$, the
surviving emigrants $E^{i}_{e,s,x,t}$ and the surviving internal
out-migrants $M^{ir}_{e,s,x,t}$ from zone $i$ to the REW are
substracted. Survivng internal-in migrants from the REW $r$ to zone $i$
which is $M^{ri}_{e,s,x,t}$ and the international immigrants from the
rest of the world to zone $i$ which is $I^{i}_{e,s,x,t}$ are then added
to the equation to obtain the population at the end interval.

@eq-projection-accounting provides the overall accounting framework of
the projection. However, many of those flows cannot be observed directly
for every age, sex, ethnic group and area. Therefore, these flows are
derived by applying the relevant rates and probabilities to the
population at risk. Except international migration, these components
entered the model directly as flows which will be discussed in later
sections. The estimation of other components are described in the
following sections.

## Births

Following @wohland2010ethnic, births are calculated at the start of the
interval so that newborn can enter the model immediately and processed
with other components. Births are generated from the female population
of reproductive age and therefore the **period-cohort** fertility rate
for each ethnic group $e$ at reproductive age $x$ in zone $i$ is given
as:

$$
f^{c,i}_{e,x,t}
=
0.5
\left(
f^{p,i}_{e,x,t}
+
f^{p,i}_{e,x+1,t}
\right)
$$ {#eq-cohort-fertility}

where the period-cohort fertility rate $f^{c,i}_{e,x,t}$ at age $x$ is
the average of the period-age fertility rate $f^{p,i}_{e,x,t}$ at age
$x$ and $f^{p,i}_{e,x+1,t}$ at $x+1$. The period-cohort fertility rates
are then applied to the female population aged 15 to 49 to calculate
births to mothers in each ethnic group $e$:

$$
B^{i}_{e,t}
=
\sum_{x=15}^{49}
f^{c,i}_{e,x,t}
P^{start,i}_{e,s,x,t},
\qquad s = Female
$$ {#eq-projected-births}

where $P^{start,i}_{e,s,x,t}$ is the female population ($s=Female$) of
ethnic group $e$ at maternal age $x$ at the start of the projection
interval. However, the ethnic group of newborn may not be necessarily
the same as mother's ethnicity when the father of the child is of a
different ethnicity. A mother-to-child ethnic transition matrix is then
applied to $B^{i}_{e,t}$ to distribute the newborn's ethnic groups. The
construction of the matrix is described in @sec-birth-transition.
Finally, the sex proportion of the projected births are 0.513 for males
and 0.487 for females. The proportions are assumed to be constant
throughout the projection period.

## Survivors and non-survivors

Mortality rates describe the level of mortality experienced by a
population but CCM requires the probability that an individual survive
from the one projection interval to the next. Following
@wohland2010ethnic, non-survivorship probabilities are derived from the
life table which is presented in @sec-mortality which is
$q^{i}_{e,s,x,t}$ and therfore, the corresponding survivorship
probability is:

$$
s^{i}_{e,s,x,t}
=
1-q^{i}_{e,s,x,t}
$$ {#eq-survivorship}

Applying these survivorship and non-survivorship probabilities to the
start population yields total survivors $TS^{i}_{e,s,x,t}$ and deaths
$D^{i}_{e,s,x,t}$ for period-cohort $x$, ethnic group $e$ and sex $s$ in
zone $i$:

$$
TS^{i}_{e,s,x,t}
=
s^{i}_{e,s,x,t}
P^{start,i}_{e,s,x,t}
$$ {#eq-total-survivors}

$$
D^{i}_{e,s,x,t}
=
q^{i}_{e,s,x,t}
P^{start,i}_{e,s,x,t}
$$ {#eq-deaths}

After the application of survivorship and non-survivorship
probabilities, survivors are aged forward by one year. For Birmingham
implementation, this is carried out before migration so that the age and
year of the population at risk is aligned with the migration inputs for
the projeciton year. This followed ONS practice to calculate migration
to the aged-on civilian population [@ons2025snppmethodology], and
departs from @wohland2010ethnic.

## Migration of survivors

After ageing, international emigration is substrated from the total
survivors $TS^{i}_{e,s,x,t}$ in Birmingham population. Unlike
@wohland2010ethnic, the B-EF projection uses direct flows of emigration
from ONS SNPP is used, no survivorship adjustment is applied. The
emigration-rate scenarios described below retain this projection timing
and convert these flows into population-dependent emigration rates.

International emigration differs according to the projection scenario.
Let $E^{i,c}_{e,s,x,t}$ denote the emigration applied under scenario
$c$, where
$c \in \{\mathrm{B\text{-}EF},\mathrm{B\text{-}ER2022}, \mathrm{LT\text{-}ER2030}\}$.

**For the B-EF scenario**:

$$
E^{i,\mathrm{B\text{-}EF}}_{e,s,x,t}
=
E^{i,\mathrm{flow}}_{e,s,x,t}
$$ {#eq-emigration-bef}

**For the B-ER2022 scenario**:

$$
E^{i,\mathrm{B\text{-}ER2022}}_{e,s,x,t}
=
r^{i,2022}_{e,s,x}
TS^{i}_{e,s,x,t},
\qquad t \geq 2022
$$ {#eq-emigration-ber2022}

The ethnic, age, sex emigration rates are calculated using the first
projection year. The rate is then held constant throughout the
projection period: $$
E^{i,\mathrm{ER2022}}_{e,s,x,t}
=
r^{i,2022}_{e,s,x}
TS^{i}_{e,s,x,t},
\qquad t \geq 2022
$$ {#eq-emigration-er2022}

**For the LT-ER2030 scenario**:

$$
E^{i,\mathrm{LT\text{-}ER2030}}_{e,s,x,t}
=
\begin{cases}
E^{i,\mathrm{flow}}_{e,s,x,t},
& t < 2030, \\[6pt]
r^{i,2030}_{e,s,x}
TS^{i}_{e,s,x,t},
& t \geq 2030.
\end{cases}
$$ {#eq-emigration-lter2030}

The emigration flows are used until 2029. The ethnic, age, sex
emigration rates are then calculated using the 2030 emigration flow and
population at risk. The emigration component under ER2030 is therefore:

$$
E^{i,\mathrm{ER2030}}_{e,s,x,t}
=
\begin{cases}
E^{i,\mathrm{flow}}_{e,s,x,t},
& t < 2030, \\[6pt]
r^{i,2030}_{e,s,x}
TS^{i}_{e,s,x,t},
& t \geq 2030.
\end{cases}
$$ {#eq-emigration-er2030}

The population surviving within the country under scenario $c$ is then
given by:

$$
WS^{i,c}_{e,s,x,t}
=
TS^{i}_{e,s,x,t}
-
E^{i,c}_{e,s,x,t}
$$ {#eq-within-country-survivors}

where $WS^{i,c}_{e,s,x,t}$ is the population surviving within the
country after mortality and international emigration under projection
scenario $c$.

Internal out-migration $M^{ir,c}_{e,s,x,t}$ from zone $i$ to the REW $r$
is then calculated by multiplying the corresponding internal
out-migration probability (detailed in @sec-internal-migration)
$m^{ir, c}_{e,s,x,t}$ to population surviving within the country
$WS^{i, c}_{e,s,x,t}$ is given as:

$$
M^{ir,c}_{e,s,x,t}
=
m^{ir}_{e,s,x,t}
WS^{i,c}_{e,s,x,t}
$$ {#eq-internal-out}

Internal in-migration $M^{ri}_{e,s,x,t}$ from REW $r$ to zone $i$ is
calculated as:

$$
M^{ri}_{e,s,x,t}
=
m^{ri}_{e,s,x,t}
P^{REW}_{e,s,x,t}
$$ {#eq-internal-in}

where $P^{REW}_{e,s,x,t}$ is the ONS projected REW population stock used
in place of a the within country survivor population for REW. The full
explanation is described in @sec-internal-migration.

## Final population

The components in @eq-projection-accounting can now be substituted with
the component-specific equations defined above. The final population
under projection scenario $c$ is given by:

$$
\begin{aligned}
P^{end,i,c}_{e,s,x,t}
={}&
\underbrace{
P^{start,i}_{e,s,x,t}
-
q^{i}_{e,s,x,t}P^{start,i}_{e,s,x,t}
-
E^{i,c}_{e,s,x,t}
}_{\text{within-country survivors}} \\[6pt]
&-
\underbrace{
m^{ir}_{e,s,x,t}
\left(
P^{start,i}_{e,s,x,t}
-
q^{i}_{e,s,x,t}P^{start,i}_{e,s,x,t}
-
E^{i,c}_{e,s,x,t}
\right)
}_{\text{internal out-migration}} \\[6pt]
&+
\underbrace{
m^{ri}_{e,s,x,t}P^{REW}_{e,s,x,t}
}_{\text{internal in-migration}}
+
\underbrace{
I^{i}_{e,s,x,t}
}_{\text{international immigration}} .
\end{aligned}
$$ {#eq-final-pop}

With regards to the newborn cohort in each projection interval, there is
no start population. Therefore, the projected births $B^{i}_{e,t}$ are
directly used in place of $P^{start,i}_{e,s,x,t}$. This implies newborn
are exposed to international emigration, internal out-migration,
internal in-migration and international immigration on the same basis as
all other cohorts.

For the terminal cohort, survivors aged 99 and 100+ are combined into a
single open-ended 100+ group. This group is carried forward in each
subsequent projection year and accumulates survivors until the end of
the projections.

\newpage

# Data sources {#sec-data}

| Component                                           | Data source                                                                    | Reference period                              | Publicly available? | Citation                                                        |
|---------------|---------------|---------------|---------------|---------------|
| **Base population**                                 | Census 2021 population by ethnic group, sex and single year of age             | 2021                                          | Yes                 | [@ons2023rm032]                                                 |
| **Fertility**                                       | NHS Maternity Services Data Set (MSDS)                                         | 2022                                          | No                  |                                                                 |
| **Fertility**                                       | ONS registered births, Birmingham                                              | 2022                                          | Yes                 | [@ons2023birthsbyarea]                                          |
| **Newborn ethnicity**                               | Derived from NHS Maternity Services Data Set (MSDS)                            | Baseline matrix                               | No                  |                                                                 |
| **Mortality**                                       | Birmingham death registration records                                          | 2022--2024                                    | No                  |                                                                 |
| **Mortality**                                       | National life tables, ONS                                                      | 2022--2024                                    | Yes                 | [@ons2025nlt]                                                   |
| **International immigration and emigration levels** | ONS 2022-based Subnational Population Projections (SNPP), migration components | 2023--2047                                    | Yes                 | [@ons2025snpp; @ons2025snppintmig]                              |
| **International migration backfill**                | ONS Mid-Year Estimates migration components                                    | 2021--2022                                    | Yes                 | [@ons2024mye]                                                   |
| **International immigration ethnicity**             | NHS GP Flag 4 registrations for Birmingham                                     | 2013--2022                                    | No                  |                                                                 |
| **International emigration nationality**            | ONS Long-Term International Migration (LTIM)                                   | June 2022                                     | Yes                 | [@ons2022ltim]                                                  |
| **Internal migration flows**                        | Census 2021 origin-destination data (ODMG03EW_LTLA)                            | One year before Census Day to Census Day 2021 | Yes                 | [@ons2023odmg03ew; @ons2023odguide]                             |
| **Internal migration ethnicity**                    | Wohland et al. (2010) internal migration ethnic-group data                     | 2001 Census / 2000--01                        | Yes                 | [@wohland2010ethnic]                                            |
| **Internal migration age-sex pattern**              | ONS detailed internal migration Data                                           | 2021--2024                                    | Yes                 | [@ons2025intmigla]                                              |
| **REW population**                                  | Census 2021 and projected England and Wales population series                  | 2021--2047                                    | Yes                 | [@ons2023rm032; @ons2025snpp; @wg2025lapp; @statswales2025lapp] |

: Main data sources used in the Birmingham ethnic population projection
{#tbl-data-sources}

\newpage

# Fertility {#sec-fertility}

## Overview

There are no available data in the UK regarding to age specific
fertility rates (ASFRs) by ethnic group. Common methods are estimating
Total Fertility rates (TFRs) from child-women ratios, using the
own-child method from Labour Survey. This section shows how the age
specific fertility rates by ethnic group are derived from limited data
sources.

## Adjustment to official birth totals

Age-specific fertility rates (ASFRs) are constructed from the NHS
Maternity Services Data Set (MSDS), filtered to the 2022 data. Recorded
were restricted to births where mother's recorded electoral wards was
within Birmingham. Maternal age between 15 and 49 was firstly grouped
into five-year reproductive age groups from 15-19 to 45-49. There is a
3.36% of birth records with missing mother's ethnicity in 2022.
Proportion redistribution of ethnic group is used to replace the missing
mother's ethnicity with the Birmingham ethnic composition at ward level.
For each combination of ward and maternal age group, the number of
births with missing mother's ethnicity was allocated across the 12
harmonised ethnic groups corresponding to the ethnic shares of that
ward.

MSDS is an administrative maternity dataset and therefore, the number
does not match to the official number of registered births reported by
ONS, thereby creating a huge discrepancy. Using the standalone MSDS
count underestimates the fertility rates of Birmingham. The resulting
approach is to estimate the shares of births by maternal ethnicity
within each maternal age group from MSDS, then apply the shares to the
Official ONS registered births for Birmingham. This preserve the age and
ethnic distribution of births but also matching the ONS count.

## Calculation of five-year ASFRs

ASFRs were initially calculated for seven five-year age groups between
15 and 49 years. For each ethnic group and maternal five-year age group,
the number of births was divided by the corresponding Biringham female
population. However, direction calucaltion of fertility rates for 12
ethnic groups and 7 age groups resulting 84 combinations of ASFRs where
$x_5$ denotes the five-year maternal age group (15–19, 20–24, ...,
45–49), $B_{e,x_5,t}^{*}$ is the number of births to mothers in ethnic
group $e$ and $P^{i}_{e,s,x_5,t}$ is the corresponding female
population.

$$
f^{p,i}_{e,x_5,t}
=
\frac{
B^{*,i}_{e,x_5,t}
}{
P^{i}_{e,s,x_5,t}
},
\qquad s=F
$$ Some combinations may contain very few or no observed births even
though the underlying fertility rate is unlikely to be exactly zero. A
smoothing technique is adopted to stabilise the rates before supply to
the projection.

## Empirical Bayes smoothing for five-year ASFRs

A group-specific Empirical Bayes (EB) smoothing mode was used to
stabilise the five-year ASFRs. This method partially shrinks these
estimates towards the mean fertility rate for the corresponding
five-year age group across ethnic groups. The gamma distribution is
specified as the prior for the underlying five-year ASFRs. The model can
be written as:

$$
f^{p,i}_{e,x_5,t}
\sim
\operatorname{Gamma}
\left(
\alpha_{x_5},
\beta_{x_5}
\right)
$$

where $f_{e,x5}$ is the true five-year ASFRs for each ethnic group. The
method of moments is used to estimate the group-specific gamma
distribution's shapes $\alpha_{x_5}$ and $\beta_{x_5}$, which uses the
group-specific mean $\mu_{x_5}$ and variance $v_{x_5}$:

$$
\alpha_{x_5}
=
\frac{\mu_{x_5}^{2}}{v_{x_5}}
$$

$$
\beta_{x_5}
=
\frac{\mu_{x_5}}{v_{x_5}}
$$

Once the shapes are obtained, the EB shrinkage can be applied to the
five-year ASFRs for each ethnic group with the following formula:

$$
\hat{f}^{EB,i}_{e,x_5,t}
=
\frac{
B^{*,i}_{e,x_5,t}
+
\alpha_{x_5}
}{
P^{i}_{e,s,x_5,t}
+
\beta_{x_5}
},
\qquad s=F
$$

## Conversion to single-year ASFRs using the Hadwiger model

The EB-smoothed fertility rates are estimated for five-year maternal age
groups. However, the cohort component model operates by single year of
age. The five-year ASFRs were therefore interpolated into single-year
rates for ages 15 to 49 using a **Hadwiger fertility model**. This
parametric model of fertility is proposed by Hadwiger (1940) to graduate
and smooth ASFRs [@hadwiger1940] . Applying the three-parameters
function to UK ethnic fertility schedules can be presented as follow:

$$
f^{H}_{e}(x)
=
\frac{R_{e} H_{e}}{T_{e}}
\left( \frac{T_{e}}{x} \right)^{3/2}
\exp
\left[
- H_{e}^{2}
\left(
\frac{T_{e}}{x} + \frac{x}{T_{e}} - 2
\right)
\right]
$$

where $f^{H}_{e}(x)$ is the fitted fertility rate for ethnic group $e$
at exact maternal age $x$, and $R_{e}$, $H_{e}$ and $T_{e}$ are the
three fitted parameters. $R_{e}$ is the total fertility of each ethnic
group served as a scaling parameter and fixes the overall level of the
curve without affecting its shape; $H_{e}$ is a shape parameter
governing the concentration of childbearing; $T_{e}$ is a location
parameter, higher value implies older ages of childbearing; $H_{e}$ is a
shape parameter controlling the concentration of childbearing around the
central age. The larger the value, the more peaked schedule.

The model is fitted to every ethnic groups individually. The parameters
were fitted by minimising the sum of squared differences between the
observed EB-smoothed five-year ASFRs and the corresponding five-year
rates implied by the fitted Hadwiger schedule. The estimated single year
ASFRs for each ethnic group are shown in @fig-fertility-asfr.

```{r}
#| label: fig-fertility-asfr
#| fig-cap: "Estimated single-year age-specific fertility rates derived from five-year grouped ASFRs, by ethnic group, Birmingham 2022"
#| fig-width: 10
#| fig-height: 10.5
#| warning: false
#| message: false
stacked = read_rds(
  here::here("data", "processed", "03_Birmingham_fertility_stacked.rds"))

eth_names = c(
  WBI = "White British", WHO = "Other White", MIX = "Mixed",
  IND = "Indian",        PAK = "Pakistani",   BAN = "Bangladeshi",
  CHI = "Chinese",       OAS = "Other Asian", BLA = "Black African",
  BLC = "Black Caribbean", OBL = "Other Black", OTH = "Other ethnic group"
)

eth_order = c("WBI", "WHO", "MIX",           # White + Mixed
              "IND", "PAK", "BAN", "CHI", "OAS",  # Asian
              "BLA", "BLC", "OBL",           # Black
              "OTH")                          # Other

eth_lookup = tibble(eth_code = names(eth_names),
                    eth_name = unname(eth_names)) %>% 
  mutate(eth_name = factor(eth_name, levels = eth_names[eth_order]))



stacked %>%
  left_join(eth_lookup, by = "eth_code") %>%
  ggplot(aes(x = single_age))+
  geom_line(aes(y = fx_obs, colour = "Grouped"),
            linewidth = 1) +
  geom_line(aes(y = corrected_fx_single, colour = "Single year"),
            linewidth = 1) +
  scale_color_manual(values = c("Single year" = "#D00070",
                                "Grouped"       = "#3c3c3b"))+
  scale_x_continuous(limits = c(15,49), breaks = seq(15,49,2))+
  labs(
    title = "Estimated singe year ASFRs from five year grouped ASFRs of women \nby ethnicity in Birmingham 2022",
    x = "Single year of age",
    y = "Age-sepcific fertility rate")+
  labs(colour = "")+
  theme_bcc(base_size =15,
            gridline_x = FALSE)+
  theme(legend.position = "bottom",
        axis.text.x = element_text(size = 9, angle = 90, hjust = 1))+
  facet_wrap(~eth_name)


```

The resulting single-year schedule provides the period-age fertility
rates $f^{p,i}_{e,x,t}$ used in the cohort-component model.

## Future fertility assumptions

There is no Birmingham-specific time series data to estimate a robust
future fertility trajectory separately for each ethnic group. Therefore,
future changes of Birmingham ASFRs follow the national fertility
assumptions set by ONS,

The differences in baseline fertility between ethnic groups are held
constant and a common proportional change in fertility over time is
applied to all ethnic group. The proportional change is obtained from
the national fertility rate where 2022 TFR is the baseline. For
Birmingham, the 2022 baseline TFR of 1.66 is held constant through 2024.
This is because there is an actual observed spike in registered births
in Birmingham during 2024 but declined in 2025. From 2025 onward, the
Birmingham fertility trajectory follows the proportional change implied
by the national assumption, declining to 1.55 by 2029 and rising
slightly to 1.59 by 2049. @tbl-fertility-assumptions shows the matching
proportion change of Birmingham and ONS TFR.

| Year | ONS TFR | Birmingham TFR | Proportion of 2022 |
|------|---------|----------------|--------------------|
| 2022 | 1.45    | 1.66           | 1                  |
| 2029 | 1.38    | 1.55           | 0.952              |
| 2049 | 1.42    | 1.59           | 0.979              |

: ONS and Birmingham TFR assumptions used in the fertility projection
{#tbl-fertility-assumptions}

As shown in @tbl-ethnic-tfr,the assumed proportional change is applied
to all the ethnic groups over the projection period.

| Ethnic group | TFR 2022 | TFR 2029 | TFR 2047 |
|--------------|---------:|---------:|---------:|
| WBI          |     1.55 |     1.45 |     1.48 |
| WHO          |     1.85 |     1.73 |     1.77 |
| MIX          |     1.45 |     1.36 |     1.39 |
| IND          |     1.12 |     1.04 |     1.07 |
| PAK          |     2.21 |     2.07 |     2.12 |
| BAN          |     1.40 |     1.31 |     1.34 |
| CHI          |     0.85 |     0.79 |     0.81 |
| OAS          |     2.23 |     2.09 |     2.13 |
| BLA          |     2.00 |     1.86 |     1.91 |
| BLC          |     1.24 |     1.16 |     1.19 |
| OBL          |     2.19 |     2.04 |     2.09 |
| OTH          |     1.63 |     1.52 |     1.55 |

: Projected total fertility rates by ethnic group, Birmingham
{#tbl-ethnic-tfr}

## Mother-to-child ethnic transfer matrix {#sec-birth-transition}

The ethnicity of a newborn is not always the same as that of the mother
and it is most obvious when parents are not from the same ethnic group.
Furthermore, parents may not report their children as mixed ethnic but
rather either one of the parent's ethnic group. A direct assignment of
births to the mother's ethnicity would misclassify a proportion of
newborns. To account for this, a mother-to-child ethnic transition
matrix was applied.

```{r}
#| label: fig-mother2child-matrix
#| fig-cap: "Mother-to-child ethnic transfer matrix"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

mixing_matrix_complete = read_csv(
  here::here("data", "processed", "04_birth_transition_matrix.csv"))


mixing_matrix_complete = mixing_matrix_complete %>% 
  mutate(percentage = round(percentage*100,1),
         group = case_when(
           percentage < 1  ~ "<1%",
           percentage >= 1  & percentage < 25 ~ "1% - <25%",
           percentage >= 25 & percentage < 50 ~ "25% - <50%",
           percentage >= 50 & percentage < 80 ~ "50% - <80%",
           percentage >= 80 ~ ">=80%"
         ),
         group = factor(group, levels = c("<1%", "1% - <25%", "25% - <50%", "50% - <80%", ">=80%")),
         eth_code_mother = factor(eth_code_mother, levels = rev(c(
           "WBI","WHO","MIX","IND","PAK","BAN","CHI","OAS","BLA","BLC","OBL","OTH"))),
         eth_code_baby = factor(eth_code_baby, levels = rev(c(
           "WBI","WHO","MIX","IND","PAK","BAN","CHI","OAS","BLA","BLC","OBL","OTH"))))


ggplot(mixing_matrix_complete, aes(eth_code_baby,eth_code_mother, fill=group,label = paste0(percentage, "%")))+
  geom_tile(color = "grey90")+
  geom_text()+
  scale_fill_manual(values = c("<1%" = "white",
                               "1% - <25%"  = bcctheme::bcc_pal("purple")(5)[5],
                               "25% - <50%" = bcctheme::bcc_pal("purple")(5)[4],
                               "50% - <80%" = bcctheme::bcc_pal("purple")(5)[3],
                               ">=80%" = bcctheme::bcc_pal("purple")(5)[2])) +
  labs(x        = "12 harmonised ethnic group of newborn", 
       y        = "12 harmonised ethnic group of mother",  
       fill = "Percentage", 
       title="Mother-to-child ethnic transfer matrix",
       subtitle = "mother's ethnicity in rows and newborn's ethnicity in columns"
      ) +
  scale_x_discrete(expand=c(0,0)) +
  scale_y_discrete(expand=c(0,0)) +
  coord_fixed()+
  theme_bcc(gridline_x = F,
            gridline_y = F,
            base_size = 15)+
  theme(legend.position = "bottom")



```

The transition matrix should be interpreted by row, with each row giving
the conditional distribution of the newborn's ethnic group given the
mother's ethnic group.Row percentages therefore sum to 100%.

\newpage

# Mortality {#sec-mortality}

## Overview

The availability of recent mortality rates at single year of age, for
both sexes and for each 12 harmonised ethnic group are limited. The
closest publicly available data came from the analysis of ethnic
differences in life expectancy by ONS in which the data used span from
2011 to 2014 which are outdated for this projection. Furthermore, the
estimates produced by ONS are age-standardised rates and cannot be
broken down into single year of age rates. Mortality rates were
therefore estimated directly using recent Birmingham death registration
data.

## Data window

Death registration records for Birmingham residents were pooled across
2022 to 2024.Deaths recorded between the COVID affected years (2020 to
2021) were explicitly excluded because the inclusion of these year for
estimation will potentially introduce permananet shocks from a temporary
pandemic. This is consistent with ONS evidence that mortality patterns
began to stabilise following the acute phase of the pandemic. Recorded
were restricted to deaths where the deceased 's corresponding electoral
wards was within Birmingham.

## Empirical Bayes shrinkage of mortality rates

Similar to the five-year ASFRs mentioned in the last section, the
calculation of mortality rates for each ethnic group for both sexes can
be unstable due to small numbers. Therefore, the same group-specific EB
shrinkage procedure was applied to the five-year grouped mortality rates
were firstly estimated.This allows mortality rates based on small
populations to be shrunk towards the wider Birmingham sex-age mortality
pattern, while estimates supported by larger numbers remain closer to
their observed values. The resulting smoothed mortality rates were then
used to construct the abridged life tables.

## Construction of abridged life tables

The smoothed mortality rates were used to construct separate abridged
life tables for each ethnic group and sex. The age intervals were
defined as 0, 1-4 and from 5-9 to 85-89 and open-ended 90+ group. For
each age interval, the smoothed central mortality rate,
$m^{EB,i}_{e,s,x,t}$, was converted to the corresponding probability of
dying within the interval, $q^{i}_{e,s,x,t}$, using the standard
life-table relationship:

$$
q^{i}_{e,s,x,t}
=
\frac{
n m^{EB,i}_{e,s,x,t}
}{
1 + (n-a_x)m^{EB,i}_{e,s,x,t}
}
$$

where $n$ is the width of the age interval and $a_x$ denotes the average
number of years lived within the interval by those who die during that
interval. For five-year age groups and 1-4, deaths were assumed to occur
in the middle of the interval. Infant mortality at age 0 was treated
separately where $a_x$ was set as the conventional assumption of 0.07,
implying that infant deaths mostly occurred within the first weeks of
life rather than evenly across the year. The 90+ group was initially
specified as an open-ended interval.

## Expansion to single-year probabilities of dying

The abridged life tables provide mortality estimated for grouped age
interval. In order to produce the single-year probabilities of dying,
the abridged life tables were therefore expanded to single year of age
using a relational method. This is a simple non-parametric expanding
techniques that borrows the detailed single-year of age mortality from a
complete life table as reference [@kostaki2001expanding]. The national
life table for 2022-2024 was used as the reference mortality schedule.
For each ethnic group, sex and age interval, an adjustment factor
$K^{i}_{e,s,x,t}$ was calculated as:

$$
K^{i}_{e,s,x,t}
=
\frac{
\ln(1-{}_n q^{i}_{e,s,x,t})
}{
\displaystyle\sum_{j=0}^{n-1}
\ln(1-q^{ref}_{s,x+j})
}
$$

where ${}_n q^{i}_{e,s,x,t}$ is the probability of dying within the
grouped age interval for ethnic group $e$ and sex $s$, and
$q^{ref}_{s,x+j}$ is the corresponding single-year probability of death
from the national reference life table. After obtaining the adjustment
factor $K^{i}_{e,s,x,t}$ , single-year mortality probabilities within
each age interval for ethnic group $e$ and sex $s$ were then calculated
as:

$$
q^{i}_{e,s,x+j,t}
=
1-
\left(
1-q^{ref}_{s,x+j}
\right)^{K^{i}_{e,s,x,t}},
\qquad j=0,\ldots,n-1
$$

The relational expansion was applied to ages 1-89. Mortality at age 0
was retained from the ethnic-specific abridged life table. For the 90+
open-ended group, the Kannisto method was employed to extrapolate
mortality across the oldest ages 90 to 100 [@thatcher1998force]. The
Kannisto model fits the specified range of older age trajectory (80-89
used) and extrapolates it outward to 90-100. This analysis used the
MortCast R pacakge to fit the model and generate the extrapolated
single-year rates [@sevcikova2021mortcast]. The expanded single-year
mortality probabilities are presented for each sex and ethnic group and
compared with the corresponding probabilities from the national
reference life table in @fig-mortality.

```{r}
#| label: fig-mortality
#| fig-cap: "Baseline single-year mortality schedules by ethnic group and sex, Birmingham"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

national_lifetable = read_xlsx(here::here("data", "mortality", "UK_national_life_table_2022_24.xlsx"))


mortality_input = read_csv(here::here("data", "processed", "01_02_Birmingham_mortality_rate_0_100_projected.csv"))

#---------------------------------------------------
#base map
#---------------------------------------------------
broad_group_map = tibble::tribble(
  ~eth_code, ~ethnic_group,          ~broad_group, ~broad_colours, 
  "WBI",     "White British",        "White",           "blue",
  "WHO",     "White Other",          "White",           "blue",
  "MIX",     "Mixed",                "Mixed",           "green",
  "IND",     "Indian",               "Asian",           "orange",
  "PAK",     "Pakistani",            "Asian",           "orange",
  "BAN",     "Bangladeshi",          "Asian",           "orange",
  "CHI",     "Chinese",              "Asian",           "orange",
  "OAS",     "Other Asian",          "Asian",           "orange",
  "BLA",     "Black African",        "Black",           "purple",
  "BLC",     "Black Caribbean",      "Black",           "purple",
  "OBL",     "Other Black",          "Black",           "purple",
  "OTH",     "Other ethnic group",   "Other",            "pink"
)

#---------------------------------------------------
#add harmonised shades + factor ordering
#---------------------------------------------------
broad_group_map = broad_group_map %>% 
  group_by(broad_colours) %>% 
  mutate(shade_number = row_number(),
         harmonised_colour = bcc_pal(palette = first(broad_colours))(n() + 2)[shade_number]) %>% 
  ungroup() %>% 
  mutate(broad_colours_hex = bcc_cols(broad_colours)) %>% 
  mutate(broad_group = factor(broad_group, levels = c("White","Mixed","Asian","Black","Other")),
         eth_code = factor(eth_code, levels = c(
           "WBI","WHO","MIX","IND","PAK","BAN","CHI","OAS","BLA","BLC","OBL","OTH"))) %>% 
  arrange(broad_group, eth_code)

#---------------------------------------------------
#widen shade spacing across the palette ramp
#---------------------------------------------------
broad_group_map = broad_group_map %>% 
  group_by(broad_colours) %>% 
  mutate(shade_number = row_number(),
         harmonised_colour = bcc_pal(palette = first(broad_colours))(9)[round(seq(2, 9, length.out = n()))][shade_number]) %>% 
  ungroup()

#---------------------------------------------------
#named colour vector for scale_*_manual
#---------------------------------------------------
harmonised_colour_values = broad_group_map %>% 
  distinct(eth_code, harmonised_colour) %>% 
  mutate(eth_code = as.character(eth_code)) %>% 
  tibble::deframe()

#---------------------------------------------------
#build named colour + label vectors from the broad-group map
#---------------------------------------------------
eth_colours = setNames(broad_group_map$harmonised_colour, broad_group_map$eth_code)
eth_labels  = setNames(broad_group_map$ethnic_group,      broad_group_map$eth_code)
eth_order   = levels(broad_group_map$eth_code)

ggplot() +
  geom_line(data = mortality_input %>% 
              filter(year == 2021) %>% 
              mutate(eth_code = factor(eth_code, levels = eth_order)),
            aes(age, 1 - survival_probability, colour = eth_code,
                group = interaction(eth_code, DEC_SEX)),
            linewidth = 0.45, alpha = 0.85) +
  geom_line(data = national_lifetable %>% mutate(DEC_SEX = Gender),
            aes(age, qx, group = DEC_SEX),
            colour = "grey30", linewidth = 0.9) +
  facet_wrap(~DEC_SEX) +
  scale_y_log10(labels = scales::label_number()) +
  scale_colour_manual(values = eth_colours,
                      breaks = eth_order) +
  labs(title = "Baseline expanded single-year mortality probabilities by ethnic group \nand sex, Birmingham",
       subtitle = "Ethnic mortality probabilities pooled from 2022-2024 death registrations; \nblack line: national life table 2022-2024",
       x = "Age", y = expression(q[x]~(log~scale)),
       colour = "Ethnic group") +
  theme_bcc(base_size = 15)+
  guides(color = guide_legend(nrow = 2, byrow = TRUE))+
   theme(legend.position = "bottom")


```

It shows that the mortality probabilities closely follow the shape of
national life table while preserving the ethnic differences in survival.
The less-smooth, stepped appearance of the ethnic is an artefect of the
relational method. It is because the same adjustment factor was applied
uniformly within each of the five-year abridged band smoothly while at
each band boundary, the adjustment factor jumps to a new value causing
steps in the curve.

## Future mortality assumptions

This projection adopted a 1% long-term mortality improvement applied to
all ethnic groups, sexes and ages. The mortality improvement was
assummed to increase gradually from 0% to 1% by the 25th projection
year. Thereafter, the annual improvement rate was held constant at 1%
for the remainder of the projection period. The life expectancies
associated with 1% mortality improvement are presented in
@tbl-life-expectancy

```{r}
#| label: tbl-life-expectancy
#| tbl-cap: "Projected life expectancy at birth under 1% mortality improvement, Birmingham"
#| warning: false
#| message: false

#------------------------------------------------------
#build the life table columns
#------------------------------------------------------
life_expectancy = mortality_input %>% 
  arrange(year, eth_code, DEC_SEX, age) %>% 
  group_by(year, eth_code, DEC_SEX) %>% 
  mutate(ax = case_when(age == 0        ~ 0.07,
                        age == max(age) ~ NA_real_,
                        TRUE            ~ 0.5),
         qx = if_else(age == max(age), 1, 1 - exp(-mx_for_deaths)),
         lx = 100000 * cumprod(lag(1 - qx, default = 1)),
         dx = lx * qx,
         Lx = if_else(age == max(age), lx / mx_for_deaths, lead(lx) + ax * dx),
         Tx = rev(cumsum(rev(Lx))),
         ex = Tx / lx
         ) %>% 
  ungroup()

#------------------------------------------------------
#reshape to the wide reporting table
#------------------------------------------------------
life_expectancy_table = life_expectancy %>% 
  filter(year %in% c(2022, 2032, 2047), age == 0) %>% 
  select(eth_code, DEC_SEX, year, ex) %>% 
  pivot_wider(names_from = c(DEC_SEX, year), values_from = ex) %>% 
  mutate(eth_code = factor(eth_code, levels = c("WBI","WHO","MIX","IND","PAK","BAN","CHI","OAS","BLA","BLC","OBL","OTH"))) %>% 
  arrange(eth_code) %>% 
  select(eth_code, Female_2022, Female_2032, Female_2047, Male_2022, Male_2032, Male_2047)

#------------------------------------------------------
#format the gt table
#------------------------------------------------------
life_expectancy_table %>% 
  gt(rowname_col = "eth_code") %>% 
  tab_spanner(label = "Female", columns = c(Female_2022, Female_2032, Female_2047)) %>% 
  tab_spanner(label = "Male",   columns = c(Male_2022, Male_2032, Male_2047)) %>% 
  cols_label(Female_2022 = "2022",
             Female_2032 = "2032",
             Female_2047 = "2047",
             Male_2022   = "2022",
             Male_2032   = "2032",
             Male_2047   = "2047"
             ) %>% 
  tab_stubhead(label = "Ethnic group") %>% 
  fmt_number(columns = everything(), decimals = 1)


```

\newpage

# International migration {#sec-immigration}

## Overview

International migration data is one of the most difficult demographic
components to estimate reliably. It is more problematic when migration
data has to be broken down into ethnicity, age and sex. There are no
reliable official sources that provide a complete record of people
entering and leaving the country at granular level required by this
project. International immigration and emigration were therefore
modelled as exogenous components using a combination of data sources.
For each migration flow, the overall annual level, age and sex schedule
and ethnic composition were estimated separately and combined
manipulatively to produce the final migration counts. This approach is
consistent with @wohland2010ethnic, with contemporary sources
substituted at each position.

Both immigration and emigration enter the projection engine as flows
rather than rates. It is unavoidable because immigrants originate
outside the projected population and no population at-risk exists to
form a denominator. Emigration can be estimated as rates and flows but
this project adopts flow instead. Given the limited data available to
estimate robust ethnicity-specific emigration rates, borrowing the ONS
SNPP emigration flow was considered the more defensive and conservative
assumption.

## Migration levels

The ONS 2022-based SNPP international migration components were employed
and filtered to Birmingham as the annual migration levels. The data
cover years 2023 to 2047. For the two years preceding the subnational
series, 2021 and 2022, observed data rather than projected levels were
used in the model. These were incorporated as back-filled inputs with
the derived ethnic-age-sex composition presented in the following
subsections. The observed data were taken from the Mid-Year Estimates
(MYE) back-series population data set for England and Wales, covering
the period from mid-2011 to mid-2023.

## Estimating the ethnic composition of immigration

The ethnic composition of international immigration was estimated from
GP registration Hlag 4 records for Birmingham from 2013 to 2022. The
Flag4 data identify individuals registering a GP whose previous address
was outside the UK or who had spent at least three months or more living
abroad. The data does not record ethnicity but rather the country of
birth.

The country of birth was redistributed to firstly the ONS 20 categories
of ethnic group using the ONS custom data webpage, and then converted to
the 12 harmonised ethnic group used in this project. Given that the
available Flag 4 data cover multiple years, the long run ethnic
composition convergence can be imposed with assumptions rather than
holding it constant. The assumptions and calculation will be presented
in section 7.6.

The derived ethnic composition was then multiplied to the ONS SNPP
immigration component at a age-sex structure. This produces an annual
series of international immigration flows by ethnicity, age and sex.

![Immigration
Flow](fig/Sec_7.3_immigration_flow.svg){#fig-immigration-flow}

@fig-immigration-flow shows the major steps in deriving the final output
of annual immigration flows by ethnicity, age and sex with the input
data including Flag4 and ONS SNPP Immigration.

## Estimating the ethnic composition of emigration

The estimation of the ethnic, age and sex compositions of emigration was
handled differently and indirectly, following the "creative" approach as
described and used by @wohland2010ethnic.

The emigration level was first divided into British and non-British
nationally using the national outflow proportion calculated from the
Long-Term International Migration series for June 2022. The same
proportion was applied to every year, age and sex. This purpose if
splitting British and non-British is to allow the two streams to carry
different ethnic compositions, since British and non-British emigration
differ substantially in composition.

To be consistent with @wohland2010ethnic, the ethnic composition of the
non-British emigration mirrors the immigration ethnic profile which is
the Flag4 data. With regards to British emigration, the Birmingham's
internal out-migration ethnic composition was assigned. The ethnic
compositions of the two streams were then held constant across the
projection period.\
\
The borrowed ethnic compositions were then multiplied to the ONS SNPP
emigration component at a age-sex structure. This result is an annual
series of ethnic-age-sex flow of emigration.

![Emigration
Flow](fig/Sec_7.4_emigration_flow.svg){#fig-emigration-flow}

@fig-emigration-flow shows the major steps in deriving the final output
of annual emigration flows by ethnicity, age and sex with the input data
including Flag4, Birmingham's internal out-migration ethnic shares and
ONS SNPP Emigration.

## Expansion of the 90 and over group

The ONS SNPP Migration components report the oldest age category as 90
and over, which is an open-ended group. To be consistent with the
projection engine with single-year structure, the open-ended 90 and over
group was expanded to single years of age 90 to 100. The procedure was
simply dividing the group total equally across eleven single-year ages.
This serves as a placeholder and should be interpreted as a pragmatic
allocation rather than a modelled age profile as it implies the same
number of migrants at age 100 as at age 90. However, the counts at these
old ages are very small and they are unlikely to influence the projected
totals.

## Future migration assumptions

Given that the annual immigration and emigration levels were taken
directly from theONS SNPP migration components, the assumptions of
future migration follow those of the ONS migration category variant. The
ONS migration category variant is set to decline from the increased
levels of earlier projection years then is held constant to its
longer-term assumption.

As the Flag4 data spans multiple years, it is possible to project the
change in ethnic shares based on assumptions. The projected ethnic
composition of immigration follows a pre-set, long-term convergence,
after which it remains constant. The observed 2022 composition was taken
as the starting point, with a pooled 2019 to 2022 composition as the
long-term target. The convergence is modelled to complete by 2028,
implying that the shift in ethnic composition over this period is
calculated as a weighted combination of the start the target points:

$$
\pi^{I,i}_{e,t}
=
(1-\lambda_t)\pi^{I,i,\text{start}}_{e}
+
\lambda_t\pi^{I,i,\text{target}}_{e}
$$

$$
\lambda_t
=
\min\left(
\frac{t-2022}{6},
1
\right)
$$

where $\pi^{I,i}_{e,t}$ is the share of international immigrants in
ethnic group $e$ in zone $i$ in year $t$. The weight $\lambda_t$
increases from 0 in 2022 to 1 in 2028, gradually shifting the ethnic
composition from the observed 2022 distribution towards the pooled 2019
to 2022 long-term target. From 2028 onward, the target ethnic
composition was held constant.

```{r}
#| label: fig-flag4projectedtrend
#| fig-cap: "Estimated ethnic composition of international immigration"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false

flag4_plot = read_rds(
  here::here("data", "processed", "Flag4_projection_plot_data.rds"))


# broad_group_map = tibble::tribble(
#   ~eth_code, ~ethnic_group,          ~broad_group, ~broad_colours, 
#   "WBI",     "White British",        "White",           "blue",
#   "WHO",     "White Other",          "White",           "blue",
#   "MIX",     "Mixed",                "Mixed",           "green",
#   "IND",     "Indian",               "Asian",           "orange",
#   "PAK",     "Pakistani",            "Asian",           "orange",
#   "BAN",     "Bangladeshi",          "Asian",           "orange",
#   "CHI",     "Chinese",              "Asian",           "orange",
#   "OAS",     "Other Asian",          "Asian",           "orange",
#   "BLA",     "Black African",        "Black",           "purple",
#   "BLC",     "Black Caribbean",      "Black",           "purple",
#   "OBL",     "Other Black",          "Black",           "purple",
#   "OTH",     "Other ethnic group",   "Other",            "pink"
# )
# 
# #=========================================================
# #ethnic group order and fixed broad-group colours
# #=========================================================
# eth_order = c(
#   "WBI", "WHO", "MIX",
#   "IND", "PAK", "BAN", "CHI", "OAS",
#   "BLA", "BLC", "OBL",
#   "OTH"
# )
# 
# eth_colours = c(
#   "WBI" = "blue",
#   "WHO" = "blue",
#   "MIX" = "green",
#   "IND" = "orange",
#   "PAK" = "orange",
#   "BAN" = "orange",
#   "CHI" = "orange",
#   "OAS" = "orange",
#   "BLA" = "purple",
#   "BLC" = "purple",
#   "OBL" = "purple",
#   "OTH" = "pink"
# )
# 
# #=========================================================
# #ethnic group order and BCC colour shades
# #=========================================================
# eth_order = c(
#   "WBI", "WHO", "MIX",
#   "IND", "PAK", "BAN", "CHI", "OAS",
#   "BLA", "BLC", "OBL",
#   "OTH"
# )
# 
# #---------------------------------------------------------
# #shade each ethnic code within its broad group
# #---------------------------------------------------------
# broad_group_map = broad_group_map %>% 
#   mutate(eth_code = factor(eth_code, levels = eth_order)) %>% 
#   arrange(eth_code) %>% 
#   group_by(broad_group, broad_colours) %>% 
#   mutate(
#     shade_number      = row_number(),
#     harmonised_colour = bcc_pal(palette = first(broad_colours))(n() + 2)[shade_number]
#   ) %>% 
#   ungroup()
# 
# eth_colours = setNames(broad_group_map$harmonised_colour, broad_group_map$eth_code)
# 

#---------------------------------------------------------
#plot observed vs projected shares
#---------------------------------------------------------
flag4_plot %>% 
  ggplot(aes(
    x        = year,
    y        = share,
    colour   = eth_code,
    linetype = period,
    group    = interaction(eth_code, period)
  )) +
  geom_line(linewidth = 1) +
  geom_point(data = ~ filter(.x, period == "Observed"), size = 1.8) +
  geom_vline(xintercept = 2022, linetype = "dotted", linewidth = 0.6) +
  scale_colour_manual(
    values = eth_colours,
    breaks = eth_order,
    labels = broad_group_map$ethnic_group[match(eth_order, broad_group_map$eth_code)]
  ) +
  scale_linetype_manual(values = c("Observed" = "solid", "Projected" = "dashed")) +
  scale_x_continuous(breaks = 2012:2028) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title    = "Ethnic composition of international immigration, Birmingham",
    subtitle = "Observed 2013-2022 and converged by 2028",
    x        = "Year",
    y        = "Share of annual immigration inflow",
    colour   = "Ethnic group",
    linetype = NULL
  ) +
  theme_bcc(base_size = 15) +
  theme(axis.text.x = element_text(angle = 45))+
  guides(color = guide_legend(nrow = 3, byrow = TRUE))+
   theme(legend.position = "bottom")




```

@fig-flag4projectedtrend shows the observed ethnic composition of
international immigration from 2013 to 2022 as solid lines. The dashed
lines show the projected shares converging by 2028. Shares are held
constant at their 2028 values for the rest of the projected years.

\newpage

# Internal migration {#sec-internal-migration}

## Overview

A different approach was taken to obtain internal migration compared
with international migration. Unlike international migration, internal
migration probabilities were calculated and supplied to the projection
engine rather than direct flows. A bi-directional internal migration
framework was used to model migration between Birmingham and the rest of
England and Wales (REW). The model projects flows from Birmingham to the
REW and from the REW to Birmingham. In each direction of flows, the
projected number of migrants is calculated by applying relevant
migration probabilities to the population at risk.

Since the census OD data record usual residence one year before the
Census day, it implies that these observed migrants are individuals who
survived from their previous residence to census date. It provides the
opportunity to separate migration component from the effect of
mortality, such that migration probabilities can be estimated using the
population at risk.

![Internal
migration](fig/Sec_8_internal_migration_rates.svg){#fig-internalmigration-rates}

@fig-internalmigration-rates shows the major steps in deriving the final
output of annual internal in- and out-migration probabilities by
ethnicity, age and sex with the input data including Census 2021
ODMG03EW_LTLA and ONS detailed age-sex internal migration data.

## Estimation of all-age in-migration probabilities

Internal out-migration is the sum of flows with recorded Birmingham
origin and non-Birmingham destination while internal in-migration is the
sum of flows with recorded non-Birmingham origin to Birmingham
destination. International immigration is also provided in the OD data
which will be in use to obtain the population at-risk denominator. These
flows were used to estimate internal in- and out-migration
probabilities.

For internal out-migration probability, the population at risk consists
of those who were resident in Birmingham at the beginning of the
migration interval that survived within England and Wales to Census
date. Since the Census population represents the population present at
the end of interval, flows of internal-in migration and immigration were
firstly removed to identify surviving stayers that had remained in
Birmingham:

$$
SS^{i}_{e}
=
P^{end,i}_{e}
-
M^{ri}_{e}
-
I^{i}_{e}
$$

where $P^{end,i}_{e}$ is the Census 2021 Birmingham population of ethnic
group $e$ (as the end stock of that interval), $M^{ri}_{e}$ is the
Internal-in migration to Birmingham and $I^{i}_{e}$ is the international
immigration.

The surviving stayers alone do not represent the full population at
risk, because those who moved from Birmingham to REW were Birmingham
residents at the beginning of the interval and survived to census date.
The observed internal out-migrants denoted as $M^{ir}_{e}$ was added
back to the surviving stayers:

$$
WS^{i}_{e}
=
SS^{i}_{e}
+
M^{ir}_{e}
$$

The derived $WS^{i}_{e}$ represents the surviving population of ethnic
group $e$ that was resident in Birmingham at the beginning of the
interval and was therefore exposed to the risk of internal
out-migration. The all-age internal out-migration probability for ethnic
group $e$ was therefore:

$$
m^{ir}_{e}
=
\frac{
M^{ir}_{e}
}{
WS^{i}_{e}
}
$$

## Estimation of all-age out-migration probabilities

For internal in-migration probability of Birmingham (which depends on
REW), the same survivor-based style calculation in principle can be
applied. Under the bi-directional CCM design, surviving stayers in REW
would be obtained by removing internal-in migration and immigration to
REW during the interval from its end-of-interval population:

$$
SS^{r}_{e}
=
P^{REW}_{e}
-
M^{ir}_{e}
-
I^{r}_{e}
$$

where $P^{REW}_{e}$ is the Census 2021 REW population (substrated
Birmingham) of ethnic group $e$, $M^{ir}_{e}$ represents migrants from
Birmingham to REW, and $I^{r}_{e}$ represents international immigration
into REW. The full population at risk of moving from REW to Birmingham
is:

$$
WS^{r}_{e}
=
SS^{r}_{e}
+
M^{ri}_{e}
$$

The derived $WS^{r}_{e}$ represents the surviving population of ethnic
group $e$ that was resident in REW at the beginning of the interval and
would therefore form the population at risk of migrating from REW to
Birmingham. From Birmingham's perspective, this corresponds to the
population at risk of internal in-migration.

However, this analysis did not use survivor-based calculation for
internal in-migration probability of Birmingham. In order to construct
$WS^{r}_{e}$ for each projection year require annual REW deaths and REW
international migration to be modelled by ethnic, sex and single of year
age. It implies the full accounting for REW to be co-projected alongside
Birmingham, which is beyond the scope of this analysis aiming to produce
projection for Birmingham only. Therefore, the ONS projected REW
population stock $P^{REW}_{e}$ was therefore directly used as the
population at risk:

$$
m^{ri}_{e}
=
\frac{
M^{ri}_{e}
}{
P^{REW}_{e}
}
$$

The difference in choices of denominator reflects how the two
populations are handled in the model. Birmingham is projected directly
within the CCM so its surivivng population at risk can be updated by
each projection year. For REW, it is provided as an exogenous population
series.

## Ethnic propensity

The Census OD data provide flows for broad ethnic groups rather than the
more granular ethnic classification required by the projection. As a
result, the broad-group internal in- and out-migration were
disaggregated to the 12 harmonised group before calculating the
migration probabilities. Within each broad ethnic group, the
distribution across the corresponding detailed ethnic groups was
borrowed from the NEWETHPOP internal migration ethnic shares data. For
REW, the 2021 census ethnic distribution was used to derive the
population stock for the 12 harmonised ethnic groups.

## Age and sex migration schedules

The ethnic-specific probabilities estimated above represent all-age
migration probabilities for both sexes. To derive ethnic-age-sex
migration probabilities, the 2021 census age-sex internal migration
schedules were used to describe the age and sex pattern of migration.

Following @wohland2010ethnic, single-year age profiles for males and
females were converted into ratios by dividing each age-sex-specific
migration probability by the mean probability of the overall migration
profile. These ratios describe the relative migration propensity at each
sex $s$ and single year of age $x$.

For out-migration, the ratio is denoted as: $$
w^{ir}_{s,x}
=
\frac{
m^{ir}_{s,x}
}{
\bar{m}^{ir}
}
$$

For in-migration, the ratio is denoted as:

$$
w^{ri}_{s,x}
=
\frac{
m^{ri}_{s,x}
}{
\bar{m}^{ri}
}
$$ where $w^{ir}_{s,x}$ and $w^{ri}_{s,x}$ are the relative age-sex
migration ratios for out-migration and in-migration respectively. These
ratios were subsequently applied to the all-age migration probabilities.
This approach assumes the independence of OD pattern and age-sex
pattern. In other words, ethnicity determines the overall level of
migration while age-sex pattern determines how that level is distributed
across sex and single year of age.

## Adjustment for ethnic age-sex composition

Issues arose when ethnic-specific all-age probabilities were simply
multiplied by the age-sex pattern. The assumption that ethnicity and
age-sex effects were independent did not hold well in this case because
ethnic groups in Birmingham differ sharply in age structure. The
ethnic-specific all-age probabilities already reflect each group's age
composition. Applying it by an age-sex pattern that peaks at the same
ages where younger ethnic groups are over-represented would inflate the
implied flows.

To solve this, a population-weighted average of the relative age-sex
migration ratios was calculated separately for each ethnic group. For
out-migration: $$
\bar{W}^{ir}_{e}
=
\frac{
\sum_{s,x}
w^{ir}_{s,x}
P^{i}_{e,s,x}
}{
\sum_{s,x}
P^{i}_{e,s,x}
}
$$ For in-migration:

$$
\bar{W}^{ri}_{e}
=
\frac{
\sum_{s,x}
w^{ri}_{s,x}
P^{REW}_{e,s,x}
}{
\sum_{s,x}
P^{REW}_{e,s,x}
}
$$ where $\bar{W}^{ir}_{e}$ is the average relative age-sex
out-migration ratio for ethnic group $e$, weighted by its Birmingham
age-sex population. Equivalently, $\bar{W}^{ri}_{e}$ is the average
relative age-sex in-migration ratio for ethnic group $e$ weighted by the
REW age-sex population. To obtain the final standardised migration
probabilities, the all-age ethnic migration probabilities were first
divided by their corresponding population-weighted average. After that,
the ratios (relative migration propensity) were multiplied to the
fraction, resulting the standardised probabilities for internal-in and
internal-out migration used in projection engine.

For out-migration: $$
m^{ir}_{e,s,x}
=
\frac{
m^{ri}_{e}
}{
\bar{W}^{ri}_{e}
}
w^{ir}_{s,x}
$$ For in-migration: $$
m^{ri}_{e,s,x}
=
\frac{
m^{ri}_{e}
}{
\bar{W}^{ri}_{e}
}
w^{ri}_{s,x}
$$

Further adjustment around internal out-migration was applied because
some ethnic-age-sex groups produced very high out-migration
probabilities. The highest values were concentrated among younger
Chinese groups, which may reflect the high mobility of student-age
population. To avoid these extremely high probabilities in generating
implausibly large outflows, the out-migration probability was capped at
0.75.

## Re-leveling and future migration rates

The Census OD data only compare individuals’usual residence one year
before Census Day with their residence on Census Day. Therefore,
additional movements between the two points in time which should be
considered as internal in- and out-migration were not captured. The
ethnic all-age probabilities were therefore re-leveled to the
corresponding ONS 2021 migration level before being projected forward.

Observed 2022-2024 data of internal migration were used to measure the
proportional change from the 2021 level for each age and sex. In other
words, 2021 is set as the baseline, and 2022-2024 levels were expressed
relative to the baseline. The median proportional change between 2022
and 2024 was used from 2025 onward and held constant for the rest of the
projection period.

\newpage

# Summary of projection assumptions {#sec-assumptions}

| Component                                   | Projection assumption                                                                                                                                                                                                                                                                                                                |
|--------------------------|----------------------------------------------|
| **Base population**                         | The 2021 Census population was used as the projection base by ethnic group, sex and single year of age but the data is not recorded in mid year but in 21 March 2021. No adjustment was made for the gap because the differences in count is minimal.                                                                                |
| **Projection horizon**                      | Population was projected annually from 2022 to 2047. Ages were modelled from 0 to 100+, with 100+ treated as the terminal open-ended age group.                                                                                                                                                                                      |
| **Ethnic groups**                           | 12 harmonised NEWETHPOP ethnic groups were used throughout the projection.                                                                                                                                                                                                                                                           |
| **Fertility**                               | The differences in baseline fertility between ethnic groups are held constant and a common proportional change in fertility over time is applied to all ethnic group. Birmingham TFR was held at 1.66 to 2024, declined to 1.55 by 2029 and then increased gradually in line with the national fertility trajectory.                 |
| **Newborn ethnicity**                       | Projected births were distributed using a mother-to-child ethnic transition matrix derived from MSDS rather than assigning every newborn directly to the mother's ethnic group.                                                                                                                                                      |
| **Mortality**                               | The same mortality improvement assumption was applied to all ethnic groups, sexes and ages. Annual mortality improvement increases gradually from 0% to 1% by the 25th projection year and is then held at 1% per year.                                                                                                              |
| **International immigration level**         | Annual immigration levels and age-sex distributions follow the ONS 2022-based SNPP migration component. Observed estimates were used for 2021 and 2022 before the SNPP series begins.                                                                                                                                                |
| **International immigration ethnic shares** | The observed 2022 ethnic composition from GP Flag 4 was used as the starting point and was assumed to converge gradually towards the pooled 2019–2022 composition by 2028. The 2028 ethnic composition was then held constant for the rest of the projection period.                                                                 |
| **International emigration level**          | Annual emigration levels and age-sex distributions follow the ONS 2022-based SNPP migration component.                                                                                                                                                                                                                               |
| **International emigration ethnicity**      | Emigration was divided into British and non-British streams using the national LTIM proportion. Non-British emigration used the international immigration ethnic profile, while British emigration used Birmingham's internal out-migration ethnic profile. These ethnic compositions were held constant over the projection period. |
| **International migration ages 90+**        | The ONS open-ended 90+ migration group was distributed equally across ages 90–100.                                                                                                                                                                                                                                                   |
| **Internal migration ethnic differences**   | Census OD flows were used to estimate ethnic-specific all-age migration probabilities. Broad Census ethnic groups were disaggregated to the 12 harmonised groups using @wohland2010ethnic internal migration ethnic-group data.                                                                                                      |
| **Internal migration age-sex pattern**      | A common age-sex migration pattern was derived from 2021 internal migration data and combined with the ethnic-specific all-age probabilities. The probabilities were adjusted for each ethnic group's age-sex composition before applying the age-sex pattern.                                                                       |
| **Internal out-migration cap**              | Annual internal out-migration probabilities were capped at 0.75 to prevent extremely high probabilities from producing implausibly large outflows.                                                                                                                                                                                   |
| **Internal migration level**                | The ethnic all-age probabilities were re-levelled to the corresponding ONS 2021 migration level before being projected forward. Observed 2022–2024 internal migration data were then used to measure proportional changes from the 2021 baseline for each age and sex.                                                               |
| **Future internal migration**               | The median proportional change between 2022 and 2024 was used from 2025 onward and held constant for the rest of the projection period.                                                                                                                                                                                              |
| **REW population**                          | REW was treated as an exogenous companion population rather than being co-projected within the Birmingham cohort-component model. For internal in-migration, the projected REW population stock was used as the population at risk.                                                                                                  |

\newpage

# Results

## Overall population projections under alternative emigration scenarios

Across all three emigration scenarios, Birmingham’s population is
projected to increase over the projection period. The differences in
these three scenarios are minimal (@fig-X101-Overall). The projected
population by 2047 reaches approximately 1,261,990 under B-EF, compared
with 1,255,962 under LT-ER2030 and approximately 1,250,000 under
B-ER2022.

```{r}
#| label: fig-X101-Overall
#| fig-cap: "Birmingham Population Projections under Alternative \nEmigration Scenarios"
#| fig-width: 10
#| fig-height: 7
#| warning: false
#| message: false


CCM_result_list = readRDS(here::here("data", "processed", "CCM_result_list.rds"))


for (i in seq_along(CCM_result_list)){
    
  full_component_projection = CCM_result_list[[i]]$full_component_projection
  
  CCM_result_list[[i]]$annual_components = full_component_projection %>%
    group_by(year) %>%
    summarise(
      births = sum(births, na.rm = TRUE),
      deaths = sum(total_deaths, na.rm = TRUE),
      
      natural_change =births - deaths,
      
      internal_in =sum(internal_in_migrants, na.rm = TRUE),
      
      internal_out =sum(internal_out_migrants, na.rm = TRUE),
      
      net_internal =internal_in - internal_out,
      
      international_in =sum(international_immigrants, na.rm = TRUE),
      
      international_out =sum(international_emigrants, na.rm = TRUE),
      
      net_international =international_in - international_out,
      
      projected_change =natural_change + net_internal +net_international,
      
      end_population = sum(end_population, na.rm = TRUE),
      start_pop = end_population-projected_change,
      
      .groups = "drop"
    )
  
  
  CCM_result_list[[i]]$annual_components_byeth = full_component_projection %>%
    group_by(year,eth_code) %>%
    summarise(
      births = sum(births, na.rm = TRUE),
      deaths = sum(total_deaths, na.rm = TRUE),
      
      natural_change =births - deaths,
      
      internal_in =sum(internal_in_migrants, na.rm = TRUE),
      
      internal_out =sum(internal_out_migrants, na.rm = TRUE),
      
      net_internal =internal_in - internal_out,
      
      international_in =sum(international_immigrants, na.rm = TRUE),
      
      international_out =sum(international_emigrants, na.rm = TRUE),
      
      net_international =international_in - international_out,
      
      projected_change =natural_change + net_internal +net_international,
      
      end_population = sum(end_population, na.rm = TRUE),
      start_pop = end_population-projected_change,
      
      .groups = "drop"
    )
  
}


plot_results_10.1 = bind_rows(
  "B-EF"      = CCM_result_list[[1]]$annual_components,
  "B-ER2022"  = CCM_result_list[[2]]$annual_components,
  "LT-ER2030" = CCM_result_list[[3]]$annual_components,
  .id = "scenario"
) %>% 
  mutate(scenario = factor(scenario, levels = c("B-EF", "B-ER2022", "LT-ER2030")),
         year = as.character(year)) %>% 
  ggplot(aes(x=year,y=end_population, group = scenario, color = scenario))+
  geom_line(size=1)+
  scale_y_continuous(limits = c(1150000,1300000))+
  theme_bcc(base_size = 15)+
  scale_colour_bcc(palette = "multi")+
  theme( axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Birmingham Population Projections under \nAlternative Emigration Scenarios")+
  labs(y="Population",
       color = "Scenarios")+
  theme(legend.title = element_text(color = bcc_cols("black")))



plot_results_10.1 



```

<!--#might add a table here  -->

The differences arise from how international emigration is
projected.Under B-EF, a fixed number is being substracted from the
projected population for each projection year and thus, the number of
emigrants does not respond to changes in proportion to the projection
population. UIn the B-ER2022 scenario, the emigration rate is fixed at
2022 rate such that it does not capture the fall in emigrants in the
earler projection years. For LT-ER2030, earlier projection years is
projected as flow until 2030 and therefore, the reduction in earlier
years is captured while after 2030, a constant rate is used so that the
changes are in proportion to the projected population.

For the rest of the results reporting, LT-ER2030 is adopted as the
principal projection scenario. THis scenario retains the annual
emigration-flow assumptions through the earlier projection period and
only switches to the constant rate once the project year reaches 2030.
The relatively small difference between these scenarios implies that the
choice does not materially alter the overall proejction trajectory.

## Changes in Birmingham's ethnic composition

### White and non-White population

The ethnic composition of Birmingham is projected to become increasingly
ethnic-diverse over the projection period. @fig-X102-linechart and
@tbl-nonwhite-share show the projected population by White and Non-White
ethnic group in count and percentage share. Birmingham's Non-White
ethnic group represent 52.8% of the end population in 2022 and is
projected to increase to 59.6% in 2032 and 64.2% by 2047. In contrast,
the White population is projected to decline from a share of 47.2% in
2022 to 40.4% in 2032 and finally reaches 35.8% by 2047.

```{r}
#| label: fig-X102-linechart
#| fig-cap: "Projected ethnic composition of Birmingham"
#| fig-width: 10
#| fig-height: 7
#| warning: false
#| message: false


CCM_result_list[[3]]$full_component_projection%>% 
  group_by(year,eth_code) %>%
  summarise(
    births = sum(births, na.rm = TRUE),
    deaths = sum(total_deaths, na.rm = TRUE),
    
    natural_change =births - deaths,
    
    internal_in =sum(internal_in_migrants, na.rm = TRUE),
    
    internal_out =sum(internal_out_migrants, na.rm = TRUE),
    
    net_internal =internal_in - internal_out,
    
    international_in =sum(international_immigrants, na.rm = TRUE),
    
    international_out =sum(international_emigrants, na.rm = TRUE),
    
    net_international =international_in - international_out,
    
    projected_change =natural_change + net_internal +net_international,
    
    end_population = sum(end_population, na.rm = TRUE),
    start_pop = end_population-projected_change,
    
    .groups = "drop"
  ) %>% 
  mutate(nonwhite = case_when(eth_code %in% c("WBI", "WHO") ~ "White",
                              TRUE ~ "Non-White")) %>% 
  group_by(year, nonwhite) %>% 
  summarise(across(where(is.numeric), ~sum(.x, na.rm = TRUE)),
            .groups = "drop") %>% 
  mutate(year = as.character(year)) %>% 
  ggplot(aes(x=year,y=end_population, colour=nonwhite,group=nonwhite)) +
  geom_line(size=1)+
  theme_bcc(base_size = 15)+
  scale_colour_bcc(palette = "multi")+
  scale_y_continuous(limits = c(10000,900000))+
  theme( axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Projected ethnic composition of Birmingham")+
  labs(y="Population",
       color = "Ethnic group")+
  theme(legend.title = element_text(color = bcc_cols("black")))
  

```

| Ethnic group |            2022 |            2032 |            2047 |
|--------------|----------------:|----------------:|----------------:|
| Non-White    | 611,935 (52.8%) | 716,925 (59.6%) | 805,903 (64.2%) |
| White        | 547,041 (47.2%) | 485,463 (40.4%) | 450,059 (35.8%) |

: Birmingham population by broad ethnic group, count (share)
{#tbl-nonwhite-share}

@fig-X102-pyramid shows how these changes are distributed across age and
sex and how these two populations differ in age structures. In terms of
absolute population count, the Non-White population dominates in the
younger age groups in 2022 and continue to be the largest by 2047.
Conversely, it is observed that the population of younger age groups in
White group is continuously shrinking.

The percentage distributions provide a clearer comparison of the age
profile within each ethnic grouping. In 2022, Non-White population is
more concentrated among children, teenagers and young adults, whereas
White population is concentrated among older adults, particularly around
age 50 and above. This pattern remains till 2032, although younger
Non-White cohorts have moved to working ages. By 2047, the percentage
shares of age profile look similar for both ethnic groups while White
population still has a larger proportion in older ages.

Overall, the projections indicate that Birmingham is getting more
ethnically diverse, Non-White population is accompanied by the aging of
the current teenagers and younger adults and a higher proportion of
children of the Non-White population than the White population by 2047.

```{r}
#| label: fig-X102-pyramid
#| fig-cap: "Projected population by age, sex and ethnicity grouping for Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false


plot_results_10.2 = CCM_result_list[[3]]$full_component_projection%>% 
  group_by(year,age,sex,eth_code) %>%
  summarise(
    births = sum(births, na.rm = TRUE),
    deaths = sum(total_deaths, na.rm = TRUE),
    
    natural_change =births - deaths,
    
    internal_in =sum(internal_in_migrants, na.rm = TRUE),
    
    internal_out =sum(internal_out_migrants, na.rm = TRUE),
    
    net_internal =internal_in - internal_out,
    
    international_in =sum(international_immigrants, na.rm = TRUE),
    
    international_out =sum(international_emigrants, na.rm = TRUE),
    
    net_international =international_in - international_out,
    
    projected_change =natural_change + net_internal +net_international,
    
    end_population = sum(end_population, na.rm = TRUE),
    start_pop = end_population-projected_change,
    
    .groups = "drop"
  ) %>% 
  mutate(nonwhite = case_when(eth_code %in% c("WBI", "WHO") ~ "White",
                              TRUE ~ "Non-White")) %>% 
  group_by(year,age,sex, nonwhite) %>% 
  summarise(across(where(is.numeric), ~sum(.x, na.rm = TRUE)),
            .groups = "drop") %>%
  mutate(age = ifelse(age>=90, "90+",as.character(age)),
         age = factor(age, levels=c(as.character(0:89), "90+")),
         sex = factor(sex, levels=c("Male", "Female"))) %>% 
  select(year,age,sex,nonwhite,end_population) %>% 
  mutate(end_population = ifelse(sex == "Male",
                                 end_population*-1,
                                 end_population)) %>% 
  filter(year %in% c(2022,2032,2047)) %>% 
  ggplot(aes(y=age, x=end_population,fill=nonwhite))+
  geom_col(width = 0.7)+
  geom_vline(xintercept = 0, colour = "white")+
  annotate("text", x = -Inf, y = Inf, label = "Male",   hjust = -0.3, vjust = 1.2, size = 3.5)+
  annotate("text", x =  Inf, y = Inf, label = "Female", hjust =  1.3, vjust = 1.2, size = 3.5)+
  facet_wrap(~year)+
  scale_x_continuous(labels = function(x) scales::comma(abs(x)))+
  scale_y_discrete(breaks = c(as.character(seq(0, 80, 10)), "90+"))+
  scale_fill_bcc(palette = "multi")+
  ggtitle("Projected population by age and sex and ethnicity grouping for Birmingham")+
  labs(subtitle = "Projections for 2022, 2032 and 2047",
       x="population")+
  theme_bcc(base_size = 15,
            gridline_x = F)+
  theme(legend.position  = "bottom",
        plot.margin      = margin(5, 20, 5, 20),
        panel.spacing.x  = unit(1.2, "lines"),
        axis.text.x      = element_text(size = 11),
        strip.background = element_blank(),
        strip.text       = element_text(face = "bold", hjust = 0.5))



overall_profile = CCM_result_list[[3]]$full_component_projection %>% 
  group_by(year, sex,age) %>% 
  summarise(
    end_population = sum(end_population, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(
    nonwhite = "All") %>% 
  group_by(year, age, sex, nonwhite) %>% 
  summarise(end_population = sum(end_population, na.rm = TRUE), .groups = "drop") %>% 
  group_by(year, nonwhite) %>% 
  mutate(pop_share = end_population / sum(end_population, na.rm = TRUE) * 100) %>% 
  ungroup() %>% 
  mutate(
    age = ifelse(age >= 90, "90+", as.character(age)),
    age = factor(age, levels = c(as.character(0:89), "90+")),
    sex = factor(sex, levels = c("Male", "Female")),
    pop_share = ifelse(sex == "Male", pop_share * -1, pop_share)
  ) %>% 
  filter(year %in% c(2022, 2032, 2047))




plot_results_10.2_pct = CCM_result_list[[3]]$full_component_projection %>% 
  group_by(year, age, sex, eth_code) %>% 
  summarise(
    end_population = sum(end_population, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(
    nonwhite = case_when(
      eth_code %in% c("WBI", "WHO") ~ "White",
      TRUE                          ~ "Non-White"
    )
  ) %>% 
  group_by(year, age, sex, nonwhite) %>% 
  summarise(end_population = sum(end_population, na.rm = TRUE), .groups = "drop") %>% 
  group_by(year, nonwhite) %>% 
  mutate(pop_share = end_population / sum(end_population, na.rm = TRUE) * 100) %>% 
  ungroup() %>% 
  mutate(
    age = ifelse(age >= 90, "90+", as.character(age)),
    age = factor(age, levels = c(as.character(0:89), "90+")),
    sex = factor(sex, levels = c("Male", "Female")),
    pop_share = ifelse(sex == "Male", pop_share * -1, pop_share)
  ) %>% 
  filter(year %in% c(2022, 2032, 2047)) %>% 
  ggplot(aes(y = age, x = pop_share)) +
  geom_col(data = overall_profile, aes(fill = nonwhite), width = 0.7) +
  geom_step(aes(colour = nonwhite, group = interaction(nonwhite, sex)),
            orientation = "y", linewidth = 0.5) +
  annotate("text", x = -Inf, y = Inf, label = "Male", hjust = -0.3, vjust = 1.2, size = 3.5) +
  annotate("text", x = Inf, y = Inf, label = "Female", hjust = 1.3, vjust = 1.2, size = 3.5) +
  facet_wrap(~year) +
  scale_x_continuous(labels = function(x) abs(x)) +
  scale_y_discrete(breaks = c(as.character(seq(0, 80, 10)), "90+")) +
  scale_colour_bcc(palette = "multi")+
  scale_fill_manual(values = c("All" = "grey80"),labels = "All", name = NULL)+
  labs(
    x = "Share of ethnic group population (%)"
  ) +
  theme_bcc(base_size = 15, gridline_x = F) +
  theme(
    legend.position  = "bottom",
    plot.margin      = margin(5, 20, 5, 20),
    panel.spacing.x  = unit(1.2, "lines"),
    axis.text.x      = element_text(size = 11),
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", hjust = 0.5)
  )

plot_results_10.2/plot_results_10.2_pct

```

### Broad ethnic groups

The White and Non-White ethnic classification provides the first
insights of the overall change in direction. To explore this in more
detail, the projection can also be examined across the five broad ethnic
group categories.

@fig-X104-broadlinechart and @tbl-broad-eth-share show the projected
population by the five broad ethnic groups in count and percentage
share. It is observed that Mixed ethnic group maintains a relatively
stable percentage share of the total population, while Other and Black
ethnic groups record modest increases in population share over the
projection period. The most obvious changes are observed among White and
Asian ethnic groups. In 2022, White ethnic group represents 47.2% as the
largest ethnic group in Birmingham and is projected to decline steadily
over the period and reaching 35.8% in 2047. In contrast, Asian ethnic
group accounts for 32% in 2022 as the second largest ethnic group in
2022 and is projected to increase over the period and reaching 40% in
2047.\

The projected Asian population overtakes the White population at around
2028, becoming the Birmingham's largest broad ethnic group for the
remainder of the projection period. This represents one of the most
significant compositional changes identified in the projection.

```{r}
#| label: fig-X104-broadlinechart
#| fig-cap: "Projected broad ethnic composition of Birmingham"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false

options(scipen = 999)

CCM_result_list[[3]]$full_component_projection%>% 
  group_by(year,eth_code) %>%
  summarise(
    births = sum(births, na.rm = TRUE),
    deaths = sum(total_deaths, na.rm = TRUE),
    
    natural_change =births - deaths,
    
    internal_in =sum(internal_in_migrants, na.rm = TRUE),
    
    internal_out =sum(internal_out_migrants, na.rm = TRUE),
    
    net_internal =internal_in - internal_out,
    
    international_in =sum(international_immigrants, na.rm = TRUE),
    
    international_out =sum(international_emigrants, na.rm = TRUE),
    
    net_international =international_in - international_out,
    
    projected_change =natural_change + net_internal +net_international,
    
    end_population = sum(end_population, na.rm = TRUE),
    start_pop = end_population-projected_change,
    
    .groups = "drop"
  ) %>% 
  mutate( broad_eth = case_when(
    eth_code %in% c("WBI", "WHO")                         ~ "White",
    eth_code == "MIX"                                     ~ "Mixed",
    eth_code %in% c("IND", "PAK", "BAN", "CHI", "OAS")   ~ "Asian",
    eth_code %in% c("BLA", "BLC", "OBL")                 ~ "Black",
    eth_code == "OTH"                                     ~ "Other"),
    broad_eth = factor(broad_eth, levels= c("White", "Mixed", "Asian", "Black", "Other"))) %>% 
  group_by(year, broad_eth) %>% 
  summarise(across(where(is.numeric), ~sum(.x, na.rm = TRUE)),
            .groups = "drop") %>% 
  mutate(year = as.character(year)) %>% 
  ggplot(aes(x=year,y=end_population, colour=broad_eth,group=broad_eth)) +
  geom_line(size=1)+
  theme_bcc(base_size = 15)+
  scale_colour_manual(values = c("White" = "#00A9E0",
                                 "Mixed" = "#75BC22",
                                 "Asian" = "#DC582A",
                                 "Black" = "#84329B",
                                 "Other" = "#D00070")) +
  scale_y_continuous(limits = c(10000,600000))+
  theme( axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Projected broad ethnic composition of Birmingham")+
  labs(y="Population",
       color = "Ethnic group")+
  theme(legend.title = element_text(color = bcc_cols("black")))


```

| Ethnic group |            2022 |            2032 |            2047 |
|--------------|----------------:|----------------:|----------------:|
| White        | 547,041 (47.2%) | 485,463 (40.4%) | 450,059 (35.8%) |
| Mixed        |   56,087 (4.8%) |   60,627 (5.0%) |   61,741 (4.9%) |
| Asian        | 370,368 (32.0%) | 437,518 (36.4%) | 502,205 (40.0%) |
| Black        | 131,212 (11.3%) | 154,099 (12.8%) | 170,209 (13.6%) |
| Other        |   54,269 (4.7%) |   64,680 (5.4%) |   71,747 (5.7%) |

: Birmingham population by broad ethnic groups, count (share)
{#tbl-broad-eth-share}

@fig-X105-broadpyramid show how these changes are distributed across age
and sex and how these five populations differ in age structure. In terms
of absolute population count, the White and Asian populations account
for the largest numbers across much of the age distribution,
specifically White population demonates the oldr ages. The Mixed, Black
and Other ethnic groups are smaller in ppulation size. Among children,
Asian counts are comparable with or higher than White counts. The
projected pyramids show these differences becoming more pronounced over
time as the ethnic composition of Birmingham changes.

The percentage distributions provide a clearer comparison of the age
profile within each broad ethnic group. The Mixed population has a
particularly young structure concentrated in newborn and children. Asian
and Black groups also have relatively greater concentrations at younger
ages in teenagers and younger adults. This younger age structure
provides greater demographic momentum for future population growth as
larger cohorts progress into reproductive and working ages. In contrast,
White population has a particularly older structure and very low
concentration in newborn, children and teenagers.

```{r}
#| label: fig-X105-broadpyramid
#| fig-cap: "Projected population by age, sex and broad ethnic group, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false



plot_results_10.5 = 
CCM_result_list[[3]]$full_component_projection%>% 
  group_by(year,age,sex,eth_code) %>%
  summarise(
    births = sum(births, na.rm = TRUE),
    deaths = sum(total_deaths, na.rm = TRUE),
    
    natural_change =births - deaths,
    
    internal_in =sum(internal_in_migrants, na.rm = TRUE),
    
    internal_out =sum(internal_out_migrants, na.rm = TRUE),
    
    net_internal =internal_in - internal_out,
    
    international_in =sum(international_immigrants, na.rm = TRUE),
    
    international_out =sum(international_emigrants, na.rm = TRUE),
    
    net_international =international_in - international_out,
    
    projected_change =natural_change + net_internal +net_international,
    
    end_population = sum(end_population, na.rm = TRUE),
    start_pop = end_population-projected_change,
    
    .groups = "drop"
  ) %>% 
  mutate( broad_eth = case_when(
    eth_code %in% c("WBI", "WHO")                         ~ "White",
    eth_code == "MIX"                                     ~ "Mixed",
    eth_code %in% c("IND", "PAK", "BAN", "CHI", "OAS")   ~ "Asian",
    eth_code %in% c("BLA", "BLC", "OBL")                 ~ "Black",
    eth_code == "OTH"                                     ~ "Other"),
    broad_eth = factor(broad_eth, levels= c("White", "Mixed", "Asian", "Black", "Other"))) %>% 
  group_by(year,age,sex, broad_eth) %>% 
  summarise(across(where(is.numeric), ~sum(.x, na.rm = TRUE)),
            .groups = "drop") %>%
  mutate(age = ifelse(age>=90, "90+",as.character(age)),
         age = factor(age, levels=c(as.character(0:89), "90+")),
         sex = factor(sex, levels=c("Male", "Female"))) %>% 
  select(year,age,sex,broad_eth,end_population) %>% 
  mutate(end_population = ifelse(sex == "Male",
                                 end_population*-1,
                                 end_population)) %>% 
  filter(year %in% c(2022,2032,2047)) %>% 
  ggplot(aes(y=age, x=end_population,fill=broad_eth))+
  geom_col(width = 0.7)+
  geom_vline(xintercept = 0, colour = "white")+
  annotate("text", x = -Inf, y = Inf, label = "Male",   hjust = -0.3, vjust = 1.2, size = 3.5)+
  annotate("text", x =  Inf, y = Inf, label = "Female", hjust =  1.3, vjust = 1.2, size = 3.5)+
  facet_wrap(~year)+
  scale_x_continuous(labels = function(x) scales::comma(abs(x)))+
  scale_y_discrete(breaks = c(as.character(seq(0, 80, 10)), "90+"))+
  scale_fill_manual(values = c("White" = "#00A9E0",
                                 "Mixed" = "#75BC22",
                                 "Asian" = "#DC582A",
                                 "Black" = "#84329B",
                                 "Other" = "#D00070"))+
  ggtitle("Projected population by age and sex and broad ethnic group for Birmingham")+
  labs(subtitle = "Projections for 2022, 2032 and 2047",
       x="population")+
  theme_bcc(base_size = 12,
            gridline_x = F)+
  theme(legend.position  = "bottom",
        plot.margin      = margin(5, 20, 5, 20),
        panel.spacing.x  = unit(1.2, "lines"),
        strip.background = element_blank(),
        strip.text       = element_text(face = "bold", hjust = 0.5))



plot_results_10.5_pct = 

CCM_result_list[[3]]$full_component_projection %>% 
  mutate( broad_eth = case_when(
    eth_code %in% c("WBI", "WHO")                         ~ "White",
    eth_code == "MIX"                                     ~ "Mixed",
    eth_code %in% c("IND", "PAK", "BAN", "CHI", "OAS")   ~ "Asian",
    eth_code %in% c("BLA", "BLC", "OBL")                 ~ "Black",
    eth_code == "OTH"                                     ~ "Other"),
    broad_eth = factor(broad_eth, levels= c("White", "Mixed", "Asian", "Black", "Other"))) %>% 
  group_by(year, age, sex, broad_eth) %>% 
  summarise(
    end_population = sum(end_population, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  group_by(year, age, sex, broad_eth) %>% 
  summarise(end_population = sum(end_population, na.rm = TRUE), .groups = "drop") %>% 
  group_by(year, broad_eth) %>% 
  mutate(pop_share = end_population / sum(end_population, na.rm = TRUE) * 100) %>% 
  ungroup() %>% 
  mutate(
    age = ifelse(age >= 90, "90+", as.character(age)),
    age = factor(age, levels = c(as.character(0:89), "90+")),
    sex = factor(sex, levels = c("Male", "Female")),
    pop_share = ifelse(sex == "Male", pop_share * -1, pop_share)
  ) %>% 
  filter(year %in% c(2022, 2032, 2047)) %>% 
  ggplot(aes(y = age, x = pop_share)) +
  geom_col(data = overall_profile, aes(fill = nonwhite), width = 0.7) +
  geom_step(aes(colour = broad_eth, group = interaction(broad_eth, sex)),
            orientation = "y", linewidth = 0.5) +
  annotate("text", x = -Inf, y = Inf, label = "Male", hjust = -0.3, vjust = 1.2, size = 3.5) +
  annotate("text", x = Inf, y = Inf, label = "Female", hjust = 1.3, vjust = 1.2, size = 3.5) +
  facet_wrap(~year) +
  scale_x_continuous(labels = function(x) abs(x)) +
  scale_y_discrete(breaks = c(as.character(seq(0, 80, 10)), "90+")) +
  scale_colour_manual(values = c("White" = "#00A9E0",
                                 "Mixed" = "#75BC22",
                                 "Asian" = "#DC582A",
                                 "Black" = "#84329B",
                                 "Other" = "#D00070"))+
  scale_fill_manual(values = c("All" = "grey80"),labels = "All", name = NULL)+
  labs(
    x = "Share of ethnic group population (%)"
  ) +
  theme_bcc(base_size = 15, gridline_x = F) +
  theme(
    legend.position  = "bottom",
    plot.margin      = margin(5, 20, 5, 20),
    panel.spacing.x  = unit(1.2, "lines"),
    axis.text.x      = element_text(size = 11),
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", hjust = 0.5)
  )




plot_results_10.5/plot_results_10.5_pct



















```

## Detailed ethnic group projections

```{r set up for the rest}
#| warning: false
#| message: false

# #===========================================================
# # function for line plots for harmonised group
# 
# 
# # harmonised_colour_values = broad_group_map %>%
# #   distinct(eth_code, harmonised_colour) %>%
# #   mutate(eth_code = as.character(eth_code)) %>%
# #   tibble::deframe()
# # 
eth_code_order = c(
"WBI", "WHO",
"MIX",
"IND", "PAK", "BAN", "CHI", "OAS",
"BLA", "BLC", "OBL",
"OTH"
)
# 



plot_result_harmonised_line = function(eth){
  
  
  annual_components_byeth_plot = CCM_result_list[[3]]$annual_components_byeth %>%
    mutate(
      eth_code = factor(
        eth_code,
        levels = eth_code_order
      )
    ) %>% 
    filter(eth_code %in% eth) %>% 
    ggplot(aes(x=year,y=end_population, colour = eth_code,
               group = eth_code))+
    geom_line(size=0.8)+
    scale_x_continuous(breaks = seq(2022, 2047, by = 5)) +
    scale_colour_manual(
      values = harmonised_colour_values,
      breaks = eth_code_order) + 
    labs(
      title = "Projected population by harmonised ethnic group",
      x = NULL,
      y = "Population",
      colour = "Broad ethnic group"
    ) +
    theme_bcc(base_size=15)
  
  return(annual_components_byeth_plot)
  
}

#===========================================================

plot_result_harmonised_line = function(eth){
  
  
  annual_components_byeth_plot = CCM_result_list[[3]]$annual_components_byeth %>%
    mutate(
      eth_code = factor(
        eth_code,
        levels = eth_code_order
      )
    ) %>% 
    filter(eth_code %in% eth) %>% 
    ggplot(aes(x=year,y=end_population, colour = eth_code,
               group = eth_code))+
    geom_line(size=0.8)+
    scale_x_continuous(breaks = seq(2022, 2047, by = 5)) +
    scale_colour_manual(
      values = harmonised_colour_values,
      breaks = eth_code_order) + 
    labs(
      title = "Projected population by harmonised ethnic group",
      x = NULL,
      y = "Population",
      colour = "Broad ethnic group"
    ) +
    theme_bcc(base_size=15)
  
  return(annual_components_byeth_plot)
  
}








#=================================================================
#function for pct pyramid plots for harmonised group

plot_result_harmonised_pyramid = function(broad = "White",
                                          eth = c("WBI", "WHO")){


harmonised_count_pyramid = CCM_result_list[[3]]$full_component_projection%>% 
  group_by(year,age,sex,eth_code) %>%
  summarise(
    births = sum(births, na.rm = TRUE),
    deaths = sum(total_deaths, na.rm = TRUE),
    
    natural_change =births - deaths,
    
    internal_in =sum(internal_in_migrants, na.rm = TRUE),
    
    internal_out =sum(internal_out_migrants, na.rm = TRUE),
    
    net_internal =internal_in - internal_out,
    
    international_in =sum(international_immigrants, na.rm = TRUE),
    
    international_out =sum(international_emigrants, na.rm = TRUE),
    
    net_international =international_in - international_out,
    
    projected_change =natural_change + net_internal +net_international,
    
    end_population = sum(end_population, na.rm = TRUE),
    start_pop = end_population-projected_change,
    
    .groups = "drop"
  ) %>% 
  group_by(year,age,sex, eth_code) %>% 
  summarise(across(where(is.numeric), ~sum(.x, na.rm = TRUE)),
            .groups = "drop") %>%
  mutate(age = ifelse(age>=90, "90+",as.character(age)),
         age = factor(age, levels=c(as.character(0:89), "90+")),
         sex = factor(sex, levels=c("Male", "Female"))) %>% 
  select(year,age,sex,eth_code,end_population) %>% 
  mutate(end_population = ifelse(sex == "Male",
                                 end_population*-1,
                                 end_population)) %>% 
  filter(year %in% c(2022,2032,2047)) %>% 
  filter(eth_code %in% eth) %>% 
  ggplot(aes(y=age, x=end_population,fill=eth_code))+
  geom_col(width = 0.7)+
  geom_vline(xintercept = 0, colour = "white")+
  annotate("text", x = -Inf, y = Inf, label = "Male",   hjust = -0.3, vjust = 1.2, size = 3.5)+
  annotate("text", x =  Inf, y = Inf, label = "Female", hjust =  1.3, vjust = 1.2, size = 3.5)+
  facet_wrap(~year)+
  scale_x_continuous(labels = function(x) scales::comma(abs(x)))+
  scale_y_discrete(breaks = c(as.character(seq(0, 80, 10)), "90+"))+
  scale_fill_manual(
    values = harmonised_colour_values,
    breaks = eth_code_order)+
  ggtitle("Projected population by age and sex and harmonised ethnic group for Birmingham")+
  labs(subtitle = "Projections for 2022, 2032 and 2047",
       x="population")+
  theme_bcc(base_size = 12,
            gridline_x = F)+
  theme(legend.position  = "bottom",
        plot.margin      = margin(5, 20, 5, 20),
        panel.spacing.x  = unit(1.2, "lines"),
        strip.background = element_blank(),
        strip.text       = element_text(face = "bold", hjust = 0.5))



#------------------------------------------------------
# pct pyramid


overall_profile_broadeth = CCM_result_list[[3]]$full_component_projection %>%
  mutate( broad_eth = case_when(
    eth_code %in% c("WBI", "WHO")                         ~ "White",
    eth_code == "MIX"                                     ~ "Mixed",
    eth_code %in% c("IND", "PAK", "BAN", "CHI", "OAS")   ~ "Asian",
    eth_code %in% c("BLA", "BLC", "OBL")                 ~ "Black",
    eth_code == "OTH"                                     ~ "Other"),
    broad_eth = factor(broad_eth, levels= c("White", "Mixed", "Asian", "Black", "Other"))) %>%  
  group_by(year, sex,age, broad_eth) %>% 
  summarise(
    end_population = sum(end_population, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  group_by(year, age, sex, broad_eth) %>% 
  summarise(end_population = sum(end_population, na.rm = TRUE), .groups = "drop") %>% 
  group_by(year, broad_eth) %>% 
  mutate(pop_share = end_population / sum(end_population, na.rm = TRUE) * 100) %>% 
  ungroup() %>% 
  mutate(
    age = ifelse(age >= 90, "90+", as.character(age)),
    age = factor(age, levels = c(as.character(0:89), "90+")),
    sex = factor(sex, levels = c("Male", "Female")),
    pop_share = ifelse(sex == "Male", pop_share * -1, pop_share)
  ) %>% 
  filter(year %in% c(2022, 2032, 2047)) %>% 
  filter(broad_eth %in% broad)





harmonised_pct_pyramid =  CCM_result_list[[3]]$full_component_projection %>% 
  group_by(year, age, sex, eth_code) %>% 
  summarise(
    end_population = sum(end_population, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  group_by(year, age, sex, eth_code) %>% 
  summarise(end_population = sum(end_population, na.rm = TRUE), .groups = "drop") %>% 
  group_by(year, eth_code) %>% 
  mutate(pop_share = end_population / sum(end_population, na.rm = TRUE) * 100) %>% 
  ungroup() %>% 
  mutate(
    age = ifelse(age >= 90, "90+", as.character(age)),
    age = factor(age, levels = c(as.character(0:89), "90+")),
    sex = factor(sex, levels = c("Male", "Female")),
    pop_share = ifelse(sex == "Male", pop_share * -1, pop_share)
  ) %>% 
  filter(year %in% c(2022, 2032, 2047)) %>% 
  filter(eth_code %in% eth) %>% 
  ggplot(aes(y = age, x = pop_share)) +
  geom_col(data = overall_profile_broadeth %>% filter(broad_eth == broad), aes(fill = "All"), width = 0.7) +
  geom_step(aes(colour = eth_code, group = interaction(eth_code, sex)),
            orientation = "y", linewidth = 0.8) +
  annotate("text", x = -Inf, y = Inf, label = "Male", hjust = -0.3, vjust = 1.2, size = 3.5) +
  annotate("text", x = Inf, y = Inf, label = "Female", hjust = 1.3, vjust = 1.2, size = 3.5) +
  facet_wrap(~year) +
  scale_x_continuous(labels = function(x) abs(x)) +
  scale_y_discrete(breaks = c(as.character(seq(0, 80, 10)), "90+")) +
  scale_colour_manual(
    values = harmonised_colour_values,
    breaks = eth_code_order)+
  scale_fill_manual(values = c("All" = "grey80"),labels = broad, name = NULL)+
  labs(
    x = "Share of ethnic group population (%)"
  ) +
  theme_bcc(base_size = 15, gridline_x = F) +
  theme(
    legend.position  = "bottom",
    plot.margin      = margin(5, 20, 5, 20),
    panel.spacing.x  = unit(1.2, "lines"),
    axis.text.x      = element_text(size = 11),
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", hjust = 0.5)
  )

combined_harmonised_pyramid = (harmonised_count_pyramid/harmonised_pct_pyramid)

return(combined_harmonised_pyramid)

}



#=================================================================
#function for pyramid profile plots referening 2022 outline for harmonised group
#same as the interacttive docuement 



pyramid_profile = CCM_result_list[[3]]$full_population_projection %>% 
  mutate(age = ifelse(age>=90, "90+",as.character(age)),
         age = factor(age, levels=c(as.character(0:89), "90+")),
         sex = factor(sex, levels=c("Male", "Female"))) %>% 
  group_by(year,eth_code,sex,age) %>%
  summarise(population = sum(population),
            .groups = "drop") %>% 
  mutate(population = ifelse(sex == "Male", population*-1, population))


pyramid_profile_harmonised = function(select_ethnic = "WBI",
                                      slider_year = c(2022, 2032,2047)){

  #---2022 reference outline, drawn in every panel-----------
  ref_profile = pyramid_profile %>% 
    filter(eth_code == select_ethnic, year == 2022) %>% 
    select(-year)
  
  #plot title specific to harmonised ethnic group
  plot_title  = paste0("Projected change in ", select_ethnic, " population by single year of age and sex\n",
                       "Birmingham, ", "2022, 2032 and 2047")
  
pyramid_profile_plot = pyramid_profile%>% 
  filter(eth_code == select_ethnic,
         year     %in% slider_year) %>% 
  ggplot(aes(y=age, x=population,fill=sex))+
  geom_col()+
  geom_step(data = ref_profile, aes(y = age, x = population, group = sex),
            colour = "black", linewidth = 0.6, direction = "mid", orientation = "y") +
  facet_wrap(~ year)+
  scale_x_continuous(
    labels = function(x) scales::comma(abs(x)))+
  scale_y_discrete(breaks = c(as.character(seq(0, 80, 10)), "90+"))+
  scale_fill_bcc(palette = "multi")+
  labs(title = plot_title)+
  theme_bcc(base_size = 15,
            gridline_y = F)+
  theme(
    legend.position  = "bottom",
    plot.margin      = margin(5, 20, 5, 20),
    panel.spacing.x  = unit(1.2, "lines"),
    axis.text.x      = element_text(size = 11),
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", hjust = 0.5)
  )

return(pyramid_profile_plot)

}



#=================================================================
#function to plot the changes in age group count and share 


age_cols =  c("0-4"   = "#C7E9B4", "5-15"  = "#7FCDBB", "16-24" = "#41B6C4",
              "25-49" = "#1D91C0", "50-64" = "#225EA8", "65+"   = "#0C2C84")


changes_in_agegroup = CCM_result_list[[3]]$full_population_projection %>% 
  mutate(age_groups = case_when(age %in% c(0:4) ~ "0-4",
                                age %in% c(5:15)~ "5-15",
                                age %in% c(16:24) ~ "16-24",
                                age %in% c(25:49) ~ "25-49",
                                age %in% c(50:64) ~ "50-64",
                                age %in% c(65:100) ~ "65+",
                                TRUE ~ NA_character_),
         age_groups = factor(age_groups, levels = rev(c("0-4", "5-15", "16-24", "25-49", "50-64", "65+")))) %>% 
  group_by(year, eth_code,age_groups) %>% 
  summarise(population = sum(population), .groups = "drop") %>% 
  filter(year %in% c(2022:2047))


change_agegroup_func = function(eth = "WBI"){
  
  #plot title specific to harmonised ethnic group
  plot_title  = paste0("Projected change in ", eth, " population by age groups")
  
age_changes_count = changes_in_agegroup %>% 
  filter(eth_code == eth) %>% 
  ggplot(aes(x = year, y = population, colour = age_groups, group = age_groups)) +
  geom_line(linewidth = 1) +
  scale_x_continuous(breaks = seq(2022, 2047, by = 5)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values = age_cols, name = "Age group") +
  labs(title = plot_title,
       subtitle = "Birmingham, 2022 to 2047",
       x     = NULL,
       y     = "Population") +
  theme_bcc(base_size = 15)+
  theme(legend.position = "bottom")+
  guides(colour = guide_legend(nrow = 1))
  

age_changes_share = changes_in_agegroup %>% 
  filter(eth_code == eth) %>% 
  ggplot(aes(x = year, y = population, fill = age_groups, group = age_groups)) +
  geom_col(position = "fill") +
  scale_x_continuous(breaks = seq(2022, 2047, by = 5)) +
  scale_fill_manual(values = age_cols)+
  labs(
    x      = NULL,
    y      = "Population (%)",
    colour = "Broad ethnic group"
  ) +
  theme_bcc(base_size = 15)+
  theme(legend.position = "bottom")+
  guides(fill = guide_legend(nrow = 1))
  


 return(age_changes_count/age_changes_share)

}



#======================================================
#waterfall plot
#2022 to 2032 , 2032 to 2047
#======================================================



waterfallplot_data = CCM_result_list[[3]]$annual_components_byeth %>% 
  filter(year %in% 2023:2032) %>%
  group_by(eth_code) %>%
  summarise(Births       = sum(births),
            Deaths       = sum(deaths) * -1,
            Internal_in  = sum(internal_in),
            Internal_out = sum(internal_out) * -1,
            Immigration  = sum(international_in),
            Emigration   = sum(international_out) * -1,
            .groups = "drop") %>%
  pivot_longer(cols = -eth_code,
               values_to = "changes",
               names_to  = "details") %>%
  mutate(details = factor(details,
                          levels = c("Births", "Deaths",
                                     "Internal_in", "Internal_out",
                                     "Immigration", "Emigration"))) %>%
  arrange(eth_code, details) %>%
  group_by(eth_code) %>%                      # <- grouping FIRST
  mutate(
    end   = cumsum(changes),
    start = lag(end,default = 0),
    ypos  = rev(seq_len(n())), 
    next_ypos = lead(ypos,default = 0),# 6..1 within each group
    pad   = max(abs(changes)) * 0.03,
    label_x     = end + ifelse(changes > 0, pad, -pad),
    label_y     = ypos,
    label_hjust = ifelse(changes > 0, 0, 1),
    change_label = paste0(ifelse(changes > 0, "+", ""),
                          scales::comma(round(changes)))
  ) %>%
  ungroup() 






waterfallplot_3247data = CCM_result_list[[3]]$annual_components_byeth %>% 
  filter(year %in% 2033:2047) %>%
  group_by(eth_code) %>%
  summarise(Births       = sum(births),
            Deaths       = sum(deaths) * -1,
            Internal_in  = sum(internal_in),
            Internal_out = sum(internal_out) * -1,
            Immigration  = sum(international_in),
            Emigration   = sum(international_out) * -1,
            .groups = "drop") %>%
  pivot_longer(cols = -eth_code,
               values_to = "changes",
               names_to  = "details") %>%
  mutate(details = factor(details,
                          levels = c("Births", "Deaths",
                                     "Internal_in", "Internal_out",
                                     "Immigration", "Emigration"))) %>%
  arrange(eth_code, details) %>%
  group_by(eth_code) %>%                      # <- grouping FIRST
  mutate(
    end   = cumsum(changes),
    start = lag(end,default = 0),
    ypos  = rev(seq_len(n())), 
    next_ypos = lead(ypos,default = 0),# 6..1 within each group
    pad   = max(abs(changes)) * 0.03,
    label_x     = end + ifelse(changes > 0, pad, -pad),
    label_y     = ypos,
    label_hjust = ifelse(changes > 0, 0, 1),
    change_label = paste0(ifelse(changes > 0, "+", ""),
                          scales::comma(round(changes)))
  ) %>%
  ungroup() 







waterfall_func = function(eth = "WBI"){

#----------------------------------------------
#for 2022-2032
#filter it 
waterfallplot_data_ready =waterfallplot_data %>% 
  filter(eth_code == eth)


start_pop = CCM_result_list[[3]]$annual_components_byeth %>%
  filter(year == 2022, eth_code == eth) %>%
  pull(end_population)

end_pop = CCM_result_list[[3]]$annual_components_byeth %>%
  filter(year == 2032, eth_code == eth) %>%
  pull(end_population)

pct_change = round((end_pop / start_pop - 1) * 100)

net_change = end_pop - start_pop
M = max(abs(waterfallplot_data_ready$changes))
M = ceiling(M / 10^(nchar(round(M)) - 2)) * 10^(nchar(round(M)) - 2) 

#title name for 2022-2032 changes 
title_2232 = paste0("Components of Projected ", eth,  " Population Change")

#---------------------------------------------
waterfall_2232 = ggplot(waterfallplot_data_ready)+
  geom_segment(aes(x = -Inf, xend = start,
                   y = ypos , yend = ypos),
               colour = "grey80",
               linewidth = 0.5)+
  geom_segment(aes(x=start, xend=end,
                   y=ypos, yend =ypos,
                   colour = changes>0),
               arrow = arrow(length = unit(8, "pt")),
               linewidth = 1)+
  geom_segment(aes(x = end,xend = end,
                   y = ypos, yend = next_ypos),
               colour = "grey50",
               linewidth = 1,
               linetype = "dashed")+
  geom_segment(data=waterfallplot_data_ready %>% filter(details=="Emigration"),
               aes(x=end,xend=end,
                   y=1,yend=0),
               colour = "black",
               linewidth = 1)+
  geom_segment(x = 0,xend = 0,
               y = 6, yend = 6.8,
               colour = "black",
               linewidth = 1)+
  # 2022 population circle
  annotate("point",x = 0, y = 6.8,
           shape = 21, fill = "white",
           size = 4, stroke = 1) +
  # 2032 population circle
  annotate("point",x = net_change, y = 0,
           shape = 21, fill = "white",
           size = 4, stroke = 1)+
  # 2022 population label
  annotate("text", x = 0, y = 7.15,
           label = paste0("2022 population: ", 
                          scales::comma(round(start_pop))),
           size = 4.5)+
  # 2032 population label
  annotate("text", x = net_change, y = -0.5,
           label = paste0("2032 projected population: \n", 
                          scales::comma(round(end_pop)),
                          " (",pct_change, "%)"),
           size = 4.5)+
  geom_text(aes(x = label_x,y = label_y,
                label = change_label,
                hjust = label_hjust,
                colour = changes > 0),
            fontface = "bold",
            size = 4,
            show.legend = FALSE)+
  scale_y_continuous(breaks = c(6.1,5.1,4.1,3.1,2.1,1.1), labels = c("Births", "Deaths",
                                                                     "Internal in-migration",
                                                                     "Internal out-migration",
                                                                     "Immigration",
                                                                     "Emigration"),
                     limits = c(-0.6, 7.5)) +
  scale_colour_manual(values = c("TRUE" = "#2166A5",
                                 "FALSE" = "#FF5A60"),
                      guide = "none")+
  scale_x_continuous(limits = c(-M, M) * 2, labels = scales::comma)+
  labs(x="",
       y="",
       title = title_2232,
       subtitle = "Birmingham, 2022-2032")+
  theme_bcc(gridline_x = F,
            gridline_y = F,
            base_size = 12)+
  theme(axis.text.x  = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(hjust = 1),
        plot.subtitle = element_text(hjust = -0.75),
        axis.text.y = element_text(size=12))


#----------------------------------------------
#for 2022-2032
#filter it 


waterfallplot_data_ready3247 = waterfallplot_3247data %>% 
  filter(eth_code == eth)


start_pop3247 = CCM_result_list[[3]]$annual_components_byeth %>%
  filter(year == 2032, eth_code == eth) %>%
  pull(end_population)

end_pop3247 = CCM_result_list[[3]]$annual_components_byeth %>%
  filter(year == 2047, eth_code == eth) %>%
  pull(end_population)

pct_change3247 = round((end_pop3247 / start_pop3247 - 1) * 100)

net_change3247 = end_pop3247 - start_pop3247
M3247 = max(abs(waterfallplot_data_ready3247$changes))
M3247 = ceiling(M3247 / 10^(nchar(round(M3247)) - 2)) * 10^(nchar(round(M3247)) - 2) 

#title name for 2022-2032 changes 
title_3247 = paste0("Components of Projected ", eth,  " Population Change")


#---3247 plot----------------------------------------------
waterfall_3247 = ggplot(waterfallplot_data_ready3247)+
  geom_segment(aes(x = -Inf, xend = start,
                   y = ypos , yend = ypos),
               colour = "grey80",
               linewidth = 0.5)+
  geom_segment(aes(x=start, xend=end,
                   y=ypos, yend =ypos,
                   colour = changes>0),
               arrow = arrow(length = unit(8, "pt")),
               linewidth = 1)+
  geom_segment(aes(x = end,xend = end,
                   y = ypos, yend = next_ypos),
               colour = "grey50",
               linewidth = 1,
               linetype = "dashed")+
  geom_segment(data=waterfallplot_data_ready3247 %>% filter(details=="Emigration"),
               aes(x=end,xend=end,
                   y=1,yend=0),
               colour = "black",
               linewidth = 1)+
  geom_segment(x = 0,xend = 0,
               y = 6, yend = 6.8,
               colour = "black",
               linewidth = 1)+
  # 2032 population circle
  annotate("point",x = 0, y = 6.8,
           shape = 21, fill = "white",
           size = 4, stroke = 1) +
  # 2047 population circle
  annotate("point",x = net_change3247, y = 0,
           shape = 21, fill = "white",
           size = 4, stroke = 1)+
  # 2032 population label
  annotate("text", x = 0, y = 7.15,
           label = paste0("2032 population: ", 
                          scales::comma(round(start_pop3247))),
           size = 4.5)+
  # 2047 population label
  annotate("text", x = net_change3247, y = -0.5,
           label = paste0("2047 projected population: \n", 
                          scales::comma(round(end_pop3247)),
                          " (",pct_change3247, "%)"),
           size = 4.5)+
  geom_text(aes(x = label_x,y = label_y,
                label = change_label,
                hjust = label_hjust,
                colour = changes > 0),
            fontface = "bold",
            size = 4,
            show.legend = FALSE)+
  scale_y_continuous(breaks = c(6.1,5.1,4.1,3.1,2.1,1.1), labels = c("Births", "Deaths",
                                                                     "Internal in-migration",
                                                                     "Internal out-migration",
                                                                     "Immigration",
                                                                     "Emigration"),
                     limits = c(-0.6, 7.5)) +
  scale_colour_manual(values = c("TRUE" = "#2166A5",
                                 "FALSE" = "#FF5A60"),
                      guide = "none")+
  scale_x_continuous(limits = c(-M3247, M3247) * 2, labels = scales::comma)+
  labs(x="",
       y="",
       title = title_3247,
       subtitle = "Birmingham, 2032-2047")+
  theme_bcc(gridline_x = F,
            gridline_y = F,
            base_size = 12)+
  theme(axis.text.x  = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(hjust = 1),
        plot.subtitle = element_text(hjust = -0.75),
        axis.text.y = element_text(size=12))

#---return both--------------------------------------------
return(waterfall_2232+waterfall_3247)



}


```

### White ethnic groups

```{r}
#| label: fig-line-white
#| fig-cap: "Projected population by harmonised ethnic group (WBI, WHO), Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false

plot_result_harmonised_line(eth = c("WBI", "WHO"))

```

```{r}
#| label: fig-pyramid-white
#| fig-cap: "Projected population by age, sex and harmonised ethnic group (WBI, WHO), Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false

plot_result_harmonised_pyramid(broad = "White", eth = c("WBI", "WHO"))

```

#### White British and Irish (WBI)

```{r}
#| label: fig-profile-wbi
#| fig-cap: "Projected change in WBI population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "WBI")

```

```{r}
#| label: fig-agegroup-wbi
#| fig-cap: "Projected WBI population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "WBI")

```

```{r}
#| label: fig-waterfall-wbi
#| fig-cap: "Components of projected WBI population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "WBI")


```

#### White Other (WHO)

```{r}
#| label: fig-profile-who
#| fig-cap: "Projected change in WHO population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "WHO")

```

```{r}
#| label: fig-agegroup-who
#| fig-cap: "Projected WHO population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "WHO")

```

```{r}
#| label: fig-waterfall-who
#| fig-cap: "Components of projected WHO population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "WHO")


```

### Asian ethnic groups

```{r}
#| label: fig-line-asian
#| fig-cap: "Projected population by harmonised ethnic group (IND, PAK, BAN, CHI, OAS), Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false

plot_result_harmonised_line(eth = c("IND", "PAK", "BAN", "CHI", "OAS"))

```

```{r}
#| label: fig-pyramid-asian
#| fig-cap: "Projected population by age, sex and harmonised ethnic group (IND, PAK, BAN, CHI, OAS), Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false

plot_result_harmonised_pyramid(broad = "Asian", eth = c("IND", "PAK", "BAN", "CHI", "OAS"))

```

#### Indian (IND)

```{r}
#| label: fig-profile-ind
#| fig-cap: "Projected change in IND population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "IND")

```

```{r}
#| label: fig-agegroup-ind
#| fig-cap: "Projected IND population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "IND")

```

```{r}
#| label: fig-waterfall-ind
#| fig-cap: "Components of projected IND population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "IND")


```

#### Pakistani (PAK)

```{r}
#| label: fig-profile-pak
#| fig-cap: "Projected change in PAK population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "PAK")

```

```{r}
#| label: fig-agegroup-pak
#| fig-cap: "Projected PAK population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "PAK")

```

```{r}
#| label: fig-waterfall-pak
#| fig-cap: "Components of projected PAK population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "PAK")


```

#### Bangladeshi (BAN)

```{r}
#| label: fig-profile-ban
#| fig-cap: "Projected change in BAN population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "BAN")

```

```{r}
#| label: fig-agegroup-ban
#| fig-cap: "Projected BAN population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "BAN")

```

```{r}
#| label: fig-waterfall-ban
#| fig-cap: "Components of projected BAN population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "BAN")


```

#### Chinese (CHI)

```{r}
#| label: fig-profile-chi
#| fig-cap: "Projected change in CHI population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "CHI")

```

```{r}
#| label: fig-agegroup-chi
#| fig-cap: "Projected CHI population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "CHI")

```

```{r}
#| label: fig-waterfall-chi
#| fig-cap: "Components of projected CHI population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "CHI")


```

#### Other Asian (OAS)

```{r}
#| label: fig-profile-oas
#| fig-cap: "Projected change in OAS population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "OAS")

```

```{r}
#| label: fig-agegroup-oas
#| fig-cap: "Projected OAS population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "OAS")

```

```{r}
#| label: fig-waterfall-oas
#| fig-cap: "Components of projected OAS population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "OAS")


```

### Black ethnic groups

```{r}
#| label: fig-line-black
#| fig-cap: "Projected population by harmonised ethnic group (BLA, BLC, OBL), Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false

plot_result_harmonised_line(eth = c("BLA", "BLC", "OBL"))

```

```{r}
#| label: fig-pyramid-black
#| fig-cap: "Projected population by age, sex and harmonised ethnic group (BLA, BLC, OBL), Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false

plot_result_harmonised_pyramid(broad = "Black", eth = c("BLA", "BLC", "OBL"))

```

#### Black African (BLA)

```{r}
#| label: fig-profile-bla
#| fig-cap: "Projected change in BLA population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "BLA")

```

```{r}
#| label: fig-agegroup-bla
#| fig-cap: "Projected BLA population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "BLA")

```

```{r}
#| label: fig-waterfall-bla
#| fig-cap: "Components of projected BLA population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "BLA")


```

#### Black Caribbean (BLC)

```{r}
#| label: fig-profile-blc
#| fig-cap: "Projected change in BLC population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "BLC")

```

```{r}
#| label: fig-agegroup-blc
#| fig-cap: "Projected BLC population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "BLC")

```

```{r}
#| label: fig-waterfall-blc
#| fig-cap: "Components of projected BLC population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "BLC")


```

#### Other Black (OBL)

```{r}
#| label: fig-profile-obl
#| fig-cap: "Projected change in OBL population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "OBL")

```

```{r}
#| label: fig-agegroup-obl
#| fig-cap: "Projected OBL population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "OBL")

```

```{r}
#| label: fig-waterfall-obl
#| fig-cap: "Components of projected OBL population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "OBL")


```

### Mixed and Other ethnic groups

```{r}
#| label: fig-line-mixedother
#| fig-cap: "Projected population by harmonised ethnic group (MIX, OTH), Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false

plot_result_harmonised_line(eth = c("MIX", "OTH"))

```

#### Mixed (MIX)

```{r}
#| label: fig-pyramid-mix
#| fig-cap: "Projected population by age, sex and harmonised ethnic group (MIX), Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false

plot_result_harmonised_pyramid(broad = "Mixed", eth = c("MIX"))

```

```{r}
#| label: fig-profile-mix
#| fig-cap: "Projected change in MIX population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "MIX")

```

```{r}
#| label: fig-agegroup-mix
#| fig-cap: "Projected MIX population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "MIX")

```

```{r}
#| label: fig-waterfall-mix
#| fig-cap: "Components of projected MIX population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "MIX")


```

#### Other ethnic group (OTH)

```{r}
#| label: fig-pyramid-oth
#| fig-cap: "Projected population by age, sex and harmonised ethnic group (OTH), Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 9
#| warning: false
#| message: false

plot_result_harmonised_pyramid(broad = "Other", eth = c("OTH"))

```

```{r}
#| label: fig-profile-oth
#| fig-cap: "Projected change in OTH population by age and sex, Birmingham, 2022, 2032 and 2047"
#| fig-width: 10
#| fig-height: 4.5
#| warning: false
#| message: false

pyramid_profile_harmonised(select_ethnic = "OTH")

```

```{r}
#| label: fig-agegroup-oth
#| fig-cap: "Projected OTH population by age group, Birmingham, 2022 to 2047"
#| fig-width: 10
#| fig-height: 10
#| warning: false
#| message: false

change_agegroup_func(eth = "OTH")

```

```{r}
#| label: fig-waterfall-oth
#| fig-cap: "Components of projected OTH population change, Birmingham, 2022-2032 and 2032-2047"
#| fig-width: 10
#| fig-height: 8
#| warning: false
#| message: false

waterfall_func(eth = "OTH")


```

## Comparison to ONS SNPP 2022

# References {.unnumbered}

::: {#refs}
:::
